---
title: "Cleaning and preparation of the Youth Surveys"
author: "Julia Rohrer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The city of Leipzig conducted Youth Surveys in the years 2010, 2015 and 2023. 
Here, we are going to start by loading the data and taking an initial look at it.

## Let's get started with the 2010 Youth Survey
In 2010, the city of Leipzig actually conducted two parallel Youth Surveys. 
All the following information was taken from the official report: "Jugend in Leipzig -- Ergebnisse einer Befragung 2010 [07/2011]" issued by Stadt Leipzig, Amt für Statistik und Wahlen in Kooperation mit dem Amt für Jugend, Familie und Bildung.

### Sampling procedure
First, students between the ages of 12 and 17 were targeted within their respective schools (although students within the same classes who happened to be older were surveyed as well). 
Second, young people between the ages of 18 and 27 were contacted separately based on registry information, to ensure that young people who already left school were represented as well.
Only the first survey has a methodology comparable to the 2015 and 2023 Youth Surveys, so we will exclusively rely on that data and ignore the survey involving older young people.

For the Youth Survey that took part in schools, initially schools were selected, in within those schools whole classes were assessed.
The targeted inital sample size was 3000, which was distributed according to the actual distribution of students between the ages of 12 to 17 in the four types of schools included and the grade levels included (Mittelschule: grades 7 to 10, Gymnasium: grades 7 to 12, Berufs-/Fachoberschule: first and second year, Förderschule: grades 7 to 10).
The corresponding schools and classrooms were then picked, prioritising schools who had participated in earlier Youth Surveys for organisational reasons.
In total, there were 3459 students in the targeted classes.
Of these, a total of 2411 (70%) participated in the Youth Survey and can be found in our data set.
The data were collected between October 18th and November 5th 2010.

To my knowledge, no weights were generated for the official report issued by the city.

### Load and clean data
```{r read2010}
# Read 2010 data
youth2010 <- readRDS("Files/umfschule2010.rds")

# prepare some basic variables
# add year
youth2010$year <- 2010

# schooltype
# existing variable
table(youth2010$`@schultyp`)
# 0: Mittelschule, secondary school (grade 7 to 10)
# 1: Gymnasium, grammar school (grade 7 to 12)
# 2: Berufs-/Fachoberschule, vocational school (1st and 2nd year)
# 3: Förderschule, special-needs school (grade 7 to 10)

youth2010$schooltype <- NA
youth2010$schooltype[youth2010$`@schultyp` == 0] <- 1 # Mittelschule
youth2010$schooltype[youth2010$`@schultyp` == 1] <- 2 # Gymnasium
youth2010$schooltype[youth2010$`@schultyp` == 2] <- 3 # Berufs-/Fachoberschule
youth2010$schooltype[youth2010$`@schultyp` == 3] <- 4 # Förderschule

# gender
table(youth2010$geschl) # original gender variable
youth2010$gender <- NA # generate new gender variable
youth2010$gender[youth2010$geschl == 1] <- "male"
youth2010$gender[youth2010$geschl == 2] <- "female"
youth2010$gender[youth2010$geschl == 3] <- "diverse"
table(youth2010$gender)

# age
table(youth2010$alter)
youth2010$age <- youth2010$alter

# migback
# migration background was assessed by asking respondents
# whether they and their parents were born in Germany
# self born in Germany? if no, which country
table(youth2010$s23ad, youth2010$s23aa, useNA = "always")

# father born in Germany? if no, which country
table(youth2010$s23bd, youth2010$s23ba, useNA = "always")

# mother born in Germany? if no, which country
table(youth2010$s23cd, youth2010$s23ca, useNA = "always")

# always lived in Germany?
table(youth2010$s24a, useNA = "always")
# in Germany since age...
table(youth2010$s24b, useNA = "always")

youth2010$migback <- NA

# Self, father, mother born in Germany
youth2010$migback[!is.na(youth2010$s23ad) & !is.na(youth2010$s23bd) & !is.na(youth2010$s23cd)] <- "none"
# Self born in Germany, father mother not born in Germany
youth2010$migback[!is.na(youth2010$s23ad) & !is.na(youth2010$s23ba) & !is.na(youth2010$s23ca)] <- "both"
# Self born in Germany, father not born in Germany, mother born in Germany
youth2010$migback[!is.na(youth2010$s23ad) & !is.na(youth2010$s23ba) & !is.na(youth2010$s23cd)] <- "father"
# Self born in Germany, father born in Germany, mother not born in Germany
youth2010$migback[!is.na(youth2010$s23ad) & !is.na(youth2010$s23bd) & !is.na(youth2010$s23ca)] <- "mother"
# Self not born in Germany
youth2010$migback[!is.na(youth2010$s23aa)] <- "self"
# If both parents in Germany, assume self is also born in germany
youth2010$migback[!is.na(youth2010$s23bd) & !is.na(youth2010$s23cd)] <- "none"

table(youth2010$migback, useNA = "always")


# Dichotomous indicator of ASSUMED language at home
# we will need this later because in 2015, migration background was
# not assessed in the same manner
youth2010$german_at_home <- 1 # default: German at home
youth2010$german_at_home[youth2010$migback == "self"|youth2010$migback == "both"] <- 0
# here, we assume that if we know that a person was born outside of Germany or both of their parents were

# if migback is missing German at home needs to be missing
youth2010$german_at_home[is.na(youth2010$migback)] <- NA
# we also know they do not speak German at home (this assumption is of course fallible)
table(youth2010$german_at_home, useNA = "always")

# Satisfaction items
# Students were asked for their satisfaction with various things
# Here, we will recode the items so that higher values indicate more satisfaction
# For all items that are comparable between the 2010, 2015 and 2023 Youth Surveys
youth2010$satis <- 6 - youth2010$s02a
youth2010$satis_money <- 6 - youth2010$s02b
youth2010$satis_friends <- 6 - youth2010$s02c
youth2010$satis_mom <- 6 - youth2010$s02d
youth2010$satis_dad <- 6 - youth2010$s02e
youth2010$satis_leisure <- 6 - youth2010$s02g
youth2010$satis_dwell <- 6 - youth2010$s02h
youth2010$satis_grades <- 6 - youth2010$s02f

table(youth2010$gender, useNA = "always")
table(youth2010$schooltype, useNA = "always")
table(youth2010$migback, useNA = "always")
table(youth2010$satis, useNA = "always")

```

### An initial look at the data

```{r descriptives2010}
# Gender
# Absolute numbers
table(youth2010$gender)
# Proportions
round(prop.table(table(youth2010$gender)), digits = 2)

# Age
table(youth2010$age)
summary(youth2010$age)
sd(youth2010$age, na.rm = TRUE)

hist(youth2010$age)
axis(1, at = seq(min(youth2010$age, na.rm = TRUE), max(youth2010$age, na.rm = TRUE), by = 1))

# School types
table(youth2010$schooltype)
round(prop.table(table(youth2010$schooltype)), digits = 2)
# 1: Mittelschule, secondary school (grade 7 to 10)
# 2: Gymnasium, grammar school (grade 7 to 12)
# 3: Berufs-/Fachoberschule, vocational school (1st and 2nd year)
# 4: Förderschule, special-needs school (grade 7 to 10)

# Migration background
table(youth2010$migback)
round(prop.table(table(youth2010$migback)), digits = 2)

# And the assumed language at home
table(youth2010$german_at_home)
round(prop.table(table(youth2010$german_at_home)), digits = 2)
```
## 2015 Youth Survey
The next Youth Survey was conducted five years later. 
The following information was taken from the official report: "Jugend in Leipzig – Ergebnisse einer Befragung 2015 [08/2015]" issued by Stadt Leipzig, Amt für Statistik und Wahlen in Kooperation mit dem
Amt für Jugend, Familie und Bildung.

### Sampling procedure
In 2015, only one survey took place, targeting students within their respective schools.

Again, schools were initially selected, and within those schools whole classes were assessed for reasons of efficiency. 
The targeted initial sample size was again 3000, which was again distributed according to the actual distribution of students across the included types of schools and grade levels (Oberschule, formerly called Mittelschule: grades 7 to 10, Gymnasium: grades 7 to *11*, Berufs-/Fachoberschule: first, second *and third* year, Förderschule: grades 7 to 10). 
In total, 3298 students were in the targeted classes.
Of these, a total of 2255 (75%) participated in the Youth Survey and can be found in our data set.
The data were collected between May 7th and June 1st 2015, which is why the students in grade 12 at Gymnasium are missing -- they already finished their final examinations (Abitur).

To my knowledge, no weights were generated for the official report issued by the city.

### Load and clean data
```{r read2015}
# Read 2015 data
library(haven)
youth2015 <- read_spss("Files/jugend2015.sav")

# prepare some basic variables
# add year
youth2015$year <- 2015

# schooltype
# existing variable
table(youth2015$schulart)
# 1: Gymnasium
# 2: Oberschule
# 3: Berufs-/Fachoberschule
# 4: Förderschule

youth2015$schooltype <- NA
youth2015$schooltype[youth2015$schulart == 2] <- 1 # Mittelschule/Oberschule
youth2015$schooltype[youth2015$schulart == 1] <- 2 # Gymnasium
youth2015$schooltype[youth2015$schulart == 3] <- 3 # Berufs-/Fachoberschule
youth2015$schooltype[youth2015$schulart == 4] <- 4 # Förderschule

# gender
table(youth2015$geschlecht)
youth2015$gender <- NA
youth2015$gender[youth2015$geschlecht == 1] <- "male"
youth2015$gender[youth2015$geschlecht == 2] <- "female"
youth2015$gender[youth2015$geschlecht == 3] <- "diverse"
table(youth2015$gender, youth2015$geschlecht)

# age
table(youth2015$alter)
youth2015$age <- youth2015$alter

# migback
# in 2015, no detailed information on migration background was collected
# but only information on the main language spoken at home
# according to personal communications, this decision was made
# (1) to increase comparability with other youth surveys and
# (2) to ask in a more child-friendly manner
youth2015$migback <- NA

table(youth2015$f19)
youth2015$german_at_home <- ifelse(youth2015$f19 == 1, 1, 0)
table(youth2015$german_at_home)


# Satisfaction items
# Students were asked for their satisfaction with various things
# Here, we will recode the items so that higher values indicate more satisfaction
# For all items that are comparable between the 2010, 2015 and 2023 Youth Surveys
youth2015$satis <- 6 - youth2015$f02a
youth2015$satis_money <- 6 - youth2015$f02b
youth2015$satis_friends <- 6 - youth2015$f02c
youth2015$satis_mom <- 6 - youth2015$f02d
youth2015$satis_dad <- 6 - youth2015$f02e
youth2015$satis_leisure <- 6 - youth2015$f02g
youth2015$satis_dwell <- 6 - youth2015$f02h
youth2015$satis_grades <- 6 - youth2015$f02f

```

### An initial look at the data

```{r descriptives2015}
# Gender
# Absolute numbers
table(youth2015$gender)
# Proportions
round(prop.table(table(youth2015$gender)), digits = 2)

# Age
table(youth2015$age)
summary(youth2015$age)
sd(youth2015$age, na.rm = TRUE)

hist(youth2015$age)
axis(1, at = seq(min(youth2015$age, na.rm = TRUE), max(youth2015$age, na.rm = TRUE), by = 1))

# School types
table(youth2015$schooltype)
round(prop.table(table(youth2015$schooltype)), digits = 2)
# 1: Mittelschule, secondary school (grade 7 to 10)
# 2: Gymnasium, grammar school (grade 7 to 11)
# 3: Berufs-/Fachoberschule, vocational school (1st, 2nd, 3rd year)
# 4: Förderschule, special-needs school (grade 7 to 10)

# Migration background
# not assessed

# And the assumed language at home
table(youth2015$german_at_home)
round(prop.table(table(youth2015$german_at_home)), digits = 2)
```

## 2023 Youth Survey
The next Youth Survey was supposed to happen in 2020 but was delayed by 3 years due to the COVID-19 pandemic.
The following information was taken from the official report: "Jugend in Leipzig 2023" issued by Stadt Leipzig, Amt für Statistik und Wahlen/Amt für Jugend un Familie/Amt für Schule.

### Sampling procedure
In 2023, the sampling procedures were adjusted.
Instead of first selecting schools and then assessing the relevant classrooms within those schools, classrooms were drawn across schools.
This change was implemented to increase the variability in the data and improve the representativeness of the results.
Furthermore, students from special-needs schools were intentionally oversampled, and weights were generated to account for this overrepresentation.

A total of 212 classrooms across 68 schools were initially selected for the cluster sample, covering a total of 5000 students, which was anticipated to result in a sample size of 3000 students given previous experiences with response rates.
When school headmaster refused to participate (which happened for a total of 9 schools), reserve schools were contacted.
According to the official report, for pragmatic reasons there were some deviations from the initially intended sample.
A total of 3053 questionnaires was filled out of which 3 had to be removed due to nonresponse of obviously fake answers.

The data were collected between March 1st and April 28th 2023, which means that students in grade 12 at Gymnasium could be included.
Approximately 65% of students filled out the questionnaire on a tablet as opposed to on paper.


### Load and clean data
```{r read2023}
# Read 2023 data
library(haven)
youth2023 <- readRDS("Files/jugend2023.rds")

# prepare some basic variables
# add year
youth2023$year <- 2023

# schooltype
# existing variable
table(youth2023$schulart)
# 1: Beruchsfachschule
# 2: Berufsschule
# 3: Förderschule
# 4: Gymnasium
# 5: Oberschule

youth2023$schooltype <- NA
youth2023$schooltype[youth2023$schulart == 5] <- 1 # Mittelschule/Oberschule
youth2023$schooltype[youth2023$schulart == 4] <- 2 # Gymnasium
youth2023$schooltype[youth2023$schulart == 1 | youth2023$schulart == 2] <- 3 # Berufs-/Fachoberschule
youth2023$schooltype[youth2023$schulart == 3] <- 4 # Förderschule

# gender
table(youth2023$geschl)
youth2023$gender <- NA
youth2023$gender[youth2023$geschl == 1] <- "male"
youth2023$gender[youth2023$geschl == 2] <- "female"
youth2023$gender[youth2023$geschl == 3] <- "diverse"
table(youth2023$gender, youth2023$geschl)

# age
table(youth2023$alter)
youth2023$age <- youth2023$alter

# migback
# country of birth
table(youth2023$staat_geb_person) # self
table(youth2023$staat_geb_vater) # father
table(youth2023$staat_geb_mutter) # mother

youth2023$migback <- NA
# Self, father, mother born in Germany
youth2023$migback[youth2023$staat_geb_person == 1 & youth2023$staat_geb_vater == 1 & youth2023$staat_geb_mutter == 1] <- "none"
# Self born in Germany, father mother not born in Germany
youth2023$migback[youth2023$staat_geb_person == 1 & youth2023$staat_geb_vater == 2 & youth2023$staat_geb_mutter == 2] <- "both"
# Self born in Germany, father not born in Germany, mother born in Germany
youth2023$migback[youth2023$staat_geb_person == 1 & youth2023$staat_geb_vater == 2 & youth2023$staat_geb_mutter == 1] <- "father"
# Self born in Germany, father born in Germany, mother not born in Germany
youth2023$migback[youth2023$staat_geb_person == 1 & youth2023$staat_geb_vater == 1 & youth2023$staat_geb_mutter == 2] <- "mother"
# Self not born in Germany
youth2023$migback[youth2023$staat_geb_person == 2] <- "self"

table(youth2023$migback, useNA = "always")

# If both parents in Germany, assume self is also born in germany
youth2023$migback[youth2023$staat_geb_vater == 1 & youth2023$staat_geb_mutter == 1] <- "none"

# ASSUMED language at home
youth2023$german_at_home <- 1
youth2023$german_at_home[youth2023$migback == "self"|youth2023$migback == "both"] <- 0

# if migback is missing German at home needs to be missing
youth2023$german_at_home[is.na(youth2023$migback)] <- NA
# we also know they do not speak German at home (this assumption is of course fallible)
table(youth2023$german_at_home, useNA = "always")


# Satisfaction items
# Students were asked for their satisfaction with various things
# Here, we will recode the items so that higher values indicate more satisfaction
# For all items that are comparable between the 2010, 2015 and 2023 Youth Surveys
youth2023$satis <- 6 - youth2023$f02a
youth2023$satis_money <- 6 - youth2023$f02b
youth2023$satis_friends <- 6 - youth2023$f02c
youth2023$satis_mom <- 6 - youth2023$f02d
youth2023$satis_dad <- 6 - youth2023$f02e
youth2023$satis_leisure <- 6 - youth2023$f02h
youth2023$satis_dwell <- 6 - youth2023$f02i
youth2023$satis_grades <- 6 - youth2023$f02f


```

### An initial look at the data

```{r descriptives2023}
# Gender
# Absolute numbers
table(youth2023$gender)
# Proportions
round(prop.table(table(youth2023$gender)), digits = 2)

# Age
table(youth2023$age)
summary(youth2023$age)
sd(youth2023$age, na.rm = TRUE)

hist(youth2023$age)
axis(1, at = seq(min(youth2023$age, na.rm = TRUE), max(youth2023$age, na.rm = TRUE), by = 1))

# School types
table(youth2023$schooltype)
round(prop.table(table(youth2023$schooltype)), digits = 2)
# 1: Mittelschule, secondary school (grade 7 to 10)
# 2: Gymnasium, grammar school (grade 7 to 11)
# 3: Berufs-/Fachoberschule, vocational school (1st, 2nd, 3rd year)
# 4: Förderschule, special-needs school (grade 7 to 10)

# Migration background
table(youth2023$migback)
round(prop.table(table(youth2023$migback)), digits = 2)

# And the assumed language at home
table(youth2023$german_at_home)
round(prop.table(table(youth2023$german_at_home)), digits = 2)

# And here are the weights
summary(youth2023$wph)

# Reverse-engineering what went into the weights
summary(lm(wph ~ as.factor(gender)*as.factor(schooltype)*as.factor(klassenstufe), data = youth2023))$r.squared
# As far as I can tell, the weights are a function of gender, schooltype and grade level (except for some minor deviations, which may be miscodings or result from assigning gender diverse students to one of the two groups)
```

## Additional variables
### Weights
The 2023 data are supposed to be analyzed with the help of weights.
Thus, we will incorporate weights into our analysis.
This requires us to add weights to the 2010 and 2015 data as well.
For now, we will assign everybody a weight of 1, which corresponds to the analyses reported in the original report.


```{r weights}

# What do the 2023 weights add up to?
sum(youth2023$wph, na.rm = TRUE)
# That is approximately, but not exactly the number of observations with assigned weights
sum(!is.na(youth2023$wph))
# So the weights are not exactly normalized (in which case they would add up to the number of observations)
# but close enough

# Equal weights for the other years
youth2010$wph <- 1
youth2015$wph <- 1


```

### Unique classroom IDs
Generate a variable that identifies each unique classroom in the data.

```{r classroomid}
# youth2010: schule, klasse A/B/C
#table(youth2010$klasse)
# Suspicion: Klasse within Schule identifies unique classrooms
youth2010$unique_classroom <- paste0(youth2010$schule, youth2010$klasse)
#table(youth2010$unique_classroom)
length(unique(youth2010$unique_classroom))
# 114 classrooms in
#table(youth2010$schule)
length(unique(youth2010$schule)) # 38 schools
str(youth2010$schule)

youth2010$school <- paste("2011", youth2010$schule, sep = "_")
youth2010$unique_classroom <- paste("2011", youth2010$unique_classroom, sep = "_")

# youth2015
# kennung, schulart, ortsteil, Klasse
# class from 1 to 11
length(unique(youth2015$kennung))
# 114 distinct values -- classrooms
# this variable is schulart + ortsteil + klasse and should identify unique classrooms
# schulart + ortsteil should uniquely identify schools
# only exception: documentation looks like there are two Berufs/fachoberschulen
# in 02 (Zentrum-Suedost)
youth2015$school <- paste0("2015_", youth2015$ortsteil, youth2015$schulart)
length(unique(youth2015$school))
#table(youth2015$school)
youth2015$school[youth2015$school == "NANA"] <- NA
youth2015$unique_classroom <- paste(2015, youth2015$kennung, sep = "_")


# youth2023
names(youth2023)
# klassenstufe, schulart, schule_amt51, klasse_amt51, schulart
#table(youth2023$schule_amt51, useNA = "always")
#table(youth2023$klasse_amt51, useNA = "always")
youth2023$test <- paste0("2023_", youth2023$schule_amt51, youth2023$klasse_amt51)
length(unique(youth2023$test)) # 168
youth2023$unique_classroom <- youth2023$test
youth2023$school <- paste0("2023_", youth2023$schule_amt51)




```

### Source variable
In 2023, some students answered the survey on tablets.
Let's generate a source variable for the other years as well.

```{r source}
# mode variable
youth2010$source <- "paper"
youth2015$source <- "paper"

```

### Match schools across years
Generate a variable that identifies each identifies identical schools across years.

```{r school_id}
# This is a mapping I manually created to match the schools
# as even identical schools had slightly different names
# between the waves (e.g., Mittleschule was renamed to oberschule)
mapping <- read.csv("Mapping_schools.csv", sep = ";")


# 2010
mapping_2010 <- mapping[mapping$Year == 2010,]
names(mapping_2010)[names(mapping_2010) == "Code"] <- "schule"

mapping_2010 <- mapping_2010[, c("schule", "school_id")]
youth2010 <- merge(youth2010, mapping_2010, by = "schule")
# all successfully matched


# 2015
library(haven)
youth2015_school <- read_spss("Files/jugend2015Schule.sav")
mapping_2015 <- mapping[mapping$Year == 2015,]
names(mapping_2015)[names(mapping_2015) == "Code"] <- "schule"

mapping_2015 <- mapping_2015[, c("schule", "school_id")]
youth2015_school <- merge(youth2015_school, mapping_2015, by = "schule", all.x = TRUE)

youth2015_school <- youth2015_school[, c("rawid", "school_id")]
youth2015 <- merge(youth2015, youth2015_school, by = "rawid", all.x = TRUE)
rm(youth2015_school)
# Some cases here cannot be matched because the students are not in the Schule.sav
# data set

# 2023: Here, we do have a proper variable
# in the main dataset, containing all names
mapping_2023 <- mapping[mapping$Year == 2023,]
names(mapping_2023)[names(mapping_2023) == "Name"] <- "schule_amt51"
mapping_2023 <- mapping_2023[, c("schule_amt51", "school_id")]

youth2023 <- merge(youth2023, mapping_2023, by = "schule_amt51", all.x = TRUE)
table(youth2023[is.na(youth2023$school_id), "schule_amt51"])
# Two schools have not been successfully matched due to special characters in their names
# However, these wont be part of the final analyses (one is a Berufsschule,
# the other one a Förderschule), so we can just ignore this issue

```
## Additional variables: What is important in life

```{r}
#############################
# All shared importance items
#############################
# Gesetz und Ordnung respektieren
# hoher Lebensstandard
# beruflich Karriere/beruflicher Erfolg
# sich politisch engagieren
# Das Leben geniessen
# Gesundheitsbewusst leben
# Sich Umweltbewusst verhalten
# Mitspracherecht in Staat und Gesellschaft haben
# Freundschaften pflegen
# Kinder haben

imp_names <- c("imp_law", "imp_standard", "imp_career", "imp_politics", "imp_enjoy", 
               "imp_health", "imp_environment", "imp_participate", "imp_friends", "imp_kids")
imp_2010 <- c("s01a", "s01b", "s01i", "s01j", "s01k", 
               "s01n", "s01o", "s01p", "s01q", "s01s")
imp_2015 <- c("f01j", "f01a", "f01l", "f01n", "f01o", 
               "f01e", "f01d", "f01c", "f01h", "f01b")
imp_2023 <- c("f01l", "f01i", "f01m", "f01n", "f01o", 
               "f01b", "f01e", "f01d", "f01h", "f01k")

# recode so that higher equals more important
youth2010[, imp_2010] <- 6 - youth2010[, imp_2010] 
names(youth2010)[names(youth2010) %in% imp_2010] <- imp_names

youth2015[, imp_2015] <- 6 - youth2015[, imp_2015] 
names(youth2015)[names(youth2015) %in% imp_2015] <- imp_names

youth2023[, imp_2023] <- 6 - youth2023[, imp_2023] 
names(youth2023)[names(youth2023) %in% imp_2023] <- imp_names

```

## Additional variables: Problems in life

```{r}

#############################
# Problem items
#############################
# 2010, 2015, 2023
# Dauerhafte Auseinandersetzungen mit Eltern
# Grossen Aerger mit Gleichaltrigen
# Probleme wegen Scheidung/Trennung Eltern
# Finanzielle Probleme
# Schulden
# Gewalt/Schlaegereien

prob_items <- c("prob_parents", "prob_peers", "prob_divorce", 
                "prob_money", "prob_debt", "prob_future", "prob_violence")
prob_2010 <- c("s30a", "s30b", "s30c", 
                "s30d", "s30e", "s30f", "s30h")
prob_2015 <- c("f31a", "f31b", "f31c", 
                "f31d", "f31e", "f31f", "f31h")

# OH NO, they changed the response format in 2023
# OH NO, in 2010 and 2015 they asked about the last two years
# In 2023, they only asked about the last year
prob_2023 <- c("f41a", "f41b", "f41c", 
               "f41d", "f41e", "f41f", "f41h")

# rename
names(youth2010)[names(youth2010) %in% prob_2010] <- prob_items
names(youth2015)[names(youth2015) %in% prob_2015] <- prob_items
names(youth2023)[names(youth2023) %in% prob_2023] <- prob_items

table(youth2010$prob_parents)
# 0 said does not apply, 1 is YES, 2 is NO
# Recode more sensibly: 1 yes, 0 no
youth2010$prob_parents[youth2010$prob_parents == 0] <- NA
youth2010$prob_parents[youth2010$prob_parents == 2] <- 0
youth2010$prob_parents[youth2010$prob_parents == 1] <- 1

table(youth2015$prob_parents)
# 1 is yes, 2 is no
# lets recode so that 0 is no, 1 is yes
youth2015$prob_parents[youth2015$prob_parents == 2] <- 0
youth2015$prob_parents[youth2015$prob_parents == 1] <- 1

```

## Additional variables: Discrimination

```{r}

#############################
# Discrimination items
#############################

# Stadtteil
# Muttersprache (2023: Herkunft)
# Geschlecht (nicht 2023)
# Schule

disc_items_2010_2015 <- c("disc_city", "disc_language", "disc_sex", "disc_school")
disc_items_2023 <- c("disc_city", "disc_language", "disc_school")

# Identify and recode items

# 2010
disc_vars_2010 <- c("s31a", "s31b", "s31c", "s31d")
table(youth2010$s31a)
youth2010[, disc_items_2010_2015] <- 6 - youth2010[, disc_vars_2010]

# 1: sehr stark, 5: ueberhaupt nicht

# 2015
disc_vars_2015 <- c("f32a", "f32b", "f32c", "f32d")
table(youth2015$f32a)
youth2015[, disc_items_2010_2015] <- 6 - youth2015[, disc_vars_2015]

# 1: sehr stark, 5: ueberhaupt nicht

# 2023
disc_vars_2023 <- c("f45a", "f45b", "f45e")
table(youth2023$f45a)
# 1: sehr haeufig, 5: nie, 6: trifft nicht zu
# This year changed the value labels 
# and introduced a "does not apply" category that does not
# really make sense -- we will assume that it's equivalent
# with the "never" response
youth2023[, disc_vars_2023][youth2023[, disc_vars_2023] == 6] <- 5
table(youth2023$f45a)
youth2023[, disc_items_2023] <- 6 - youth2023[, disc_vars_2023]
youth2023$disc_sex <- NA

```

## Combined dataframe and inclusion criteria
Let's create one dataframe for everything we will need later.

```{r combining}
# Add unique ids
intersect(youth2010$pnum, youth2015$rawid)
intersect(youth2010$pnum, youth2023$rawid)
intersect(youth2023$rawid, youth2015$rawid)
# none of these values intersect, so we can use them
youth2010$id <- youth2010$pnum
youth2015$id <- youth2015$rawid
youth2023$id <- youth2023$rawid


vars <- c("id", "year", "wph", "schooltype", "gender", "age", "german_at_home", "migback",
          "satis", "satis_money", "satis_friends", "satis_mom", "satis_dad",
          "satis_leisure", "satis_dwell", "satis_grades", "school", "unique_classroom", "school_id", "source", 
          imp_names, prob_items, disc_items_2010_2015)
combined <- rbind(youth2010[, vars], youth2015[, vars], youth2023[, vars])


table(youth2010$gender)
table(youth2015$gender)
table(youth2023$gender)



```
To simply the research question somewhat, we will also limit ourselves to students in particular schooltypes, of certain ages.
In the German system and in Leipzig, the standard is that after primary school, students attend either Mittelschule or Gymnasium.
Special needs school target a different population that may have shifted over time, due to both changes in the criteria for special schools, and to intentional changes to the survey design (the 2023 Survey intentionally oversampled students from Förderschule).
Students who finished Mittelschule (or, less frequently, Gymnasium), may move on to Berufs-/Fachoberschule.
But this is a heterogeneous crowd, mixing people living at home and those who have moved out.
So, our analysis will focus on Mittelschule and Gymnasium -- and within those, students with typical ages




```{r inclusion}
combined_original <- combined

##################
# Numbers included in Table 1: Data set
##################
# Total n per year
table(combined_original$year)

# Classrooms per year
length(unique(combined_original$unique_classroom[combined_original$year == 2010]))
length(unique(combined_original$unique_classroom[combined_original$year == 2015]))
length(unique(combined_original$unique_classroom[combined_original$year == 2023]))

# Schools per year
length(unique(combined_original$school[combined_original$year == 2010]))
length(unique(combined_original$school[combined_original$year == 2015]))
length(unique(combined_original$school[combined_original$year == 2023]))

# Schooltype
round(prop.table(table(combined_original$schooltype[combined_original$year == 2010], useNA = "always")), 2)
round(prop.table(table(combined_original$schooltype[combined_original$year == 2015], useNA = "always")), 2)
round(prop.table(table(combined_original$schooltype[combined_original$year == 2023], useNA = "always")), 2)

# Gender
round(prop.table(table(combined_original$gender[combined_original$year == 2010], useNA = "always")), 2)
round(prop.table(table(combined_original$gender[combined_original$year == 2015], useNA = "always")), 2)
round(prop.table(table(combined_original$gender[combined_original$year == 2023], useNA = "always")), 2)

# Age
combined_original$age[combined_original$age <= 11] <- 11
combined_original$age[combined_original$age >= 19] <- 19

round(prop.table(table(combined_original$age[combined_original$year == 2010], useNA = "always")), 2)
round(prop.table(table(combined_original$age[combined_original$year == 2015], useNA = "always")), 2)
round(prop.table(table(combined_original$age[combined_original$year == 2023], useNA = "always")), 2)

# German language at home
round(prop.table(table(combined_original$german_at_home[combined_original$year == 2010], useNA = "always")), 2)

round(prop.table(table(combined_original$german_at_home[combined_original$year == 2015], useNA = "always")), 2)

round(prop.table(table(combined_original$german_at_home[combined_original$year == 2023], useNA = "always")), 2)

# Exclusion criteria/filter
# Reporting a different gender was not possible in 2010 and 2015
# We thus cannot compare the gender diverse students in 2023 to gender diverse students in earlier years
# Which is why we unfortunately have to exclude them for the present purposes
combined$gender[combined$gender == "diverse"] <- NA

# only keep known gender
combined <- combined[!is.na(combined$gender),]

# known language at home
combined <- combined[!is.na(combined$german_at_home),]

# known age
combined <- combined[!is.na(combined$age),]

# known schooltype
combined <- combined[!is.na(combined$schooltype),]

# known outcome
combined <- combined[!is.na(combined$satis),]


combined$gender <- as.factor(combined$gender)
combined$school <- as.factor(combined$school)
combined$unique_classroom <- as.factor(combined$unique_classroom)

# Limit age range
combined <- combined[combined$age >= 12 & combined$age <= 18,]

# Generate a copy with all school types that we will need for something else
combined_all_schooltypes <- combined

# Main analysis sample
combined <- combined[combined$schooltype == 1 | combined$schooltype == 2,]
table(combined$age)
hist(combined$age)

# Standardize the outcome across everyone
combined$satis_std <- scale(combined$satis)

##################
# Numbers included in Table 1: Analysis
##################
# Total n per year
table(combined$year)

# Classrooms per year
length(unique(combined$unique_classroom[combined$year == 2010]))
length(unique(combined$unique_classroom[combined$year == 2015]))
length(unique(combined$unique_classroom[combined$year == 2023]))

# Schools per year
length(unique(combined$school[combined$year == 2010]))
length(unique(combined$school[combined$year == 2015]))
length(unique(combined$school[combined$year == 2023]))

# Schooltype
round(prop.table(table(combined$schooltype[combined$year == 2010], useNA = "always")), 2)
round(prop.table(table(combined$schooltype[combined$year == 2015], useNA = "always")), 2)
round(prop.table(table(combined$schooltype[combined$year == 2023], useNA = "always")), 2)

# Gender
round(prop.table(table(combined$gender[combined$year == 2010], useNA = "always")), 2)
round(prop.table(table(combined$gender[combined$year == 2015], useNA = "always")), 2)
round(prop.table(table(combined$gender[combined$year == 2023], useNA = "always")), 2)

# Age
combined$age[combined$age <= 11] <- 11
combined$age[combined$age >= 19] <- 19

round(prop.table(table(combined$age[combined$year == 2010], useNA = "always")), 2)
round(prop.table(table(combined$age[combined$year == 2015], useNA = "always")), 2)
round(prop.table(table(combined$age[combined$year == 2023], useNA = "always")), 2)

# German language at home
round(prop.table(table(combined$german_at_home[combined$year == 2010], useNA = "always")), 2)

round(prop.table(table(combined$german_at_home[combined$year == 2015], useNA = "always")), 2)

round(prop.table(table(combined$german_at_home[combined$year == 2023], useNA = "always")), 2)

table(combined_original$school_id, combined_original$schooltype)
save.image("prep.RData")



```
