---
title: "Reproducing the explanandum and alternative satisfaction measures"
author: "Julia Rohrer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    includes: 
      before_body: nav.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The explanandum: Widening gender gap in general life satisfaction
First, we want to reproduce the finding from the press release that states that girls in particular have become unhappier.

```{r explanandum}
load("prep.RData")
library(haven)
library(marginaleffects)
library(ggplot2)

# Simple model that estimates the means of life satisfaction by gender, survey year, and their interaction
explanandum <- lm(satis ~ as.factor(year) + gender + gender:as.factor(year), 
                  data = combined)

# We can take a look at the coefficients which map onto what we are interested in in a rather straightforward manner
summary(explanandum)

# However, do get directly started, with the marginaleffects framework...
# Here we generate predicted values by gender and year
pred <- predictions(explanandum,
                 by = c("gender", "year"))


# Plot the model-implied means
# Figure 1 in the manuscript

ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum.png", width = 4, height = 3)


# Calculating the gender gap by year
# With the help of average counterfactual comparisons
# (the averaging here does not really matter given that our model does not
# allow for any variability within the groups)

comps <- avg_comparisons(explanandum, 
                         variables = "gender", # counterfactual comparisons over gender (what if boy/girl)
                         by = "year") # separately for each year

# Take a look at the resulting gender gaps
print(comps)

# Test maximum contrast (gender gap in 2010 vs gender gap in 2023)
max_comp <- avg_comparisons(explanandum, variables = "gender", by = "year",
                hypothesis = "b1 = b3") # This tests the first versus the third number in the output
print(max_comp)


# We can also compare the other gender gaps
# Compare gender gap in 2010 with gender gap in 2015
print(avg_comparisons(explanandum, variables = "gender", by = "year",
                hypothesis = "b1 = b2"))
# Compare gender gap in 2015 with gender gap in 2023
print(avg_comparisons(explanandum, variables = "gender", by = "year",
                hypothesis = "b2 = b3"))

# Express the gender gaps in SD units (SD of life satisfaction across all observations)
round(comps$estimate[1]/sd(combined$satis), 2)
round(comps$estimate[2]/sd(combined$satis), 2)
round(comps$estimate[3]/sd(combined$satis), 2)

# Express widening of gender gap in SDs 
# Denominator is SD of life satisfaction across all observations
round(max_comp$estimate/sd(combined$satis), 2)
```


## Alternative measures of satisfaction
Let's repeat the previous analyses for all satisfaction measures in our data that are available for all three years.
Here, we also generate the individual panels of Figure 2.

### Satisfaction with the amount of money with which you have to get by (financial satisfaction)

```{r money}
# Estimate means
model_money <- lm(satis_money ~ as.factor(year) + gender + gender:as.factor(year), 
                  data = combined)
summary(model_money)
pred <- predictions(model_money,
                 by = c("gender", "year"))

# Test maximum contrast
max_comp <- avg_comparisons(model_money, variables = "gender", by = "year",
                            hypothesis = "b1 = b3")
print(max_comp)
# Express in SD units
round(max_comp$estimate/sd(combined$satis_money, na.rm = TRUE), 2)

# Plot
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean financial satisfaction") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_money.png", width = 4, height = 3)
```

### Satisfaction with your relationships to friends

```{r friends}
# Estimate means
model_friends <- lm(satis_friends ~ as.factor(year) + gender + gender:as.factor(year), 
                  data = combined)
summary(model_friends)

pred <- predictions(model_friends,
                 by = c("gender", "year"))
# Test maximum contrast
max_comp <- avg_comparisons(model_friends, variables = "gender", by = "year",
                            hypothesis = "b1 = b3")
print(max_comp)

# Express in SD units
round(max_comp$estimate/sd(combined$satis_friends, na.rm = TRUE), 2)

# Plot
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean friendship satisfaction") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_friends.png", width = 4, height = 3)
```

### Satisfaction with your relationship to your mother

```{r mom}
# Estimate means
model_mom <- lm(satis_mom ~ as.factor(year) + gender + gender:as.factor(year), 
                  data = combined)
summary(model_mom)
pred <- predictions(model_mom,
                 by = c("gender", "year"))
# Test maximum contrast
max_comp <- avg_comparisons(model_mom, variables = "gender", by = "year",
                            hypothesis = "b1 = b3")
print(max_comp)

# Express in SD units
round(max_comp$estimate/sd(combined$satis_mom, na.rm = TRUE), 2)

# Plot
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean satisfaction w mother") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_mom.png", width = 4, height = 3)
```

### Satisfaction with your relationship to your father

```{r dad}
# Estimate means
model_dad <- lm(satis_dad ~ as.factor(year) + gender + gender:as.factor(year), 
                  data = combined)
summary(model_dad)
pred <- predictions(model_dad,
                 by = c("gender", "year"))

# Test maximum contrast
max_comp <- avg_comparisons(model_dad, variables = "gender", by = "year",
                            hypothesis = "b1 = b3")
print(max_comp)

# Express in SD units
round(max_comp$estimate/sd(combined$satis_dad, na.rm = TRUE), 2)

# Plot
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean satisfaction w father") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_dad.png", width = 4, height = 3)
```

### Satisfaction with your leisute time activities

```{r leisure}
# Estimate means
model_leisure <- lm(satis_leisure ~ as.factor(year) + gender + gender:as.factor(year), 
                  data = combined)
summary(model_leisure)
pred <- predictions(model_leisure,
                 by = c("gender", "year"))

# Test maximum contrast
max_comp <- avg_comparisons(model_leisure, variables = "gender", by = "year",
                            hypothesis = "b1 = b3")
print(max_comp)

# Express in SD units
round(max_comp$estimate/sd(combined$satis_leisure, na.rm = TRUE), 2)

# Plot
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean leisure satisfaction") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_leisure.png", width = 4, height = 3)
```

### Satisfaction with your school grades

```{r grades}
# Estimate means
model_grades <- lm(satis_grades ~ as.factor(year) + gender + gender:as.factor(year), 
                  data = combined)
summary(model_grades)
pred <- predictions(model_grades,
                 by = c("gender", "year"))

# Test maximum contrast
max_comp <- avg_comparisons(model_grades, variables = "gender", by = "year",
                            hypothesis = "b1 = b3")
print(max_comp)

# Express in SD units
round(max_comp$estimate/sd(combined$satis_grades, na.rm = TRUE), 2)

# Plot
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean school grades satisfaction") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_grades.png", width = 4, height = 3)
```

### Satisfaction with your housing situation

```{r dwell}
# Estimate means
model_dwell <- lm(satis_dwell ~ as.factor(year) + gender + gender:as.factor(year), 
                  data = combined)
summary(model_dwell)
pred <- predictions(model_dwell,
                 by = c("gender", "year"))

# Test maximum contrast
max_comp <- avg_comparisons(model_dwell, variables = "gender", by = "year",
                            hypothesis = "b1 = b3")
print(max_comp)

# Express in SD units
round(max_comp$estimate/sd(combined$satis_dwell, na.rm = TRUE), 2)

# Plot
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean housing satisfaction") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_dwell.png", width = 4, height = 3)
```