---
title: "Add demographic variables"
author: "Julia Rohrer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    includes: 
      before_body: nav.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Add demographic variables
Here, we are going to add in demographic variables in an iterative manner, also looking at their associations with the outcome to gauge whether they could plausibly explain away anything.
If you just want to see the results reported in the manuscript, you can immediately skip to the final model at the bottom of the page.

## Check for differences between the years
### Age

```{r diffs_age}
load("prep.RData")
library(marginaleffects)
library(ggplot2)
library(patchwork)
library(splines)

# Are there age differences between the years?
age_diffs <- lm(age ~ as.factor(year)*gender, data = combined)

print(predictions(age_diffs,
                 by = c("gender", "year")))
print(avg_comparisons(age_diffs))
print(avg_comparisons(age_diffs, variable = "gender", by = "year"))

```

### Schooltype 

```{r diffs_schooltype}
# Recode to turn it into a Gymnasium dummy
combined$gymnasium <- ifelse(combined$schooltype == 2, 1, 0)

# Are there age differences between the years?
# Fit a simple linear model to estimate probabilities
schooltype_diffs <- lm(gymnasium ~ as.factor(year)*gender, data = combined)
# Differences between years
print(predictions(schooltype_diffs,
                 by = c("year"), vcov = "HC3"))
# Differences between genders
print(predictions(schooltype_diffs,
                 by = c("gender"), vcov = "HC3"))
# Differences between genders and years
print(predictions(schooltype_diffs,
                 by = c("gender", "year"), vcov = "HC3"))

# Gender gap by year
print(avg_comparisons(schooltype_diffs, variable = "gender", by = "year", vcov = "HC3"))
```


```{r diffs_migback}
# Are the differences in migration background between the years?

# Simple linear probability model predicting a migration background
mig_lang_diffs <- lm(mig_lang ~ as.factor(year)*gender, data = combined)

# Means by survey year
print(predictions(mig_lang_diffs,
                 by = c("year"), vcov = "HC3"))

# Means by gender
print(predictions(mig_lang_diffs,
                 by = c("gender"), vcov = "HC3"))

# Means by gender and year
print(predictions(mig_lang_diffs,
                 by = c("gender", "year"), vcov = "HC3"))

# Gender gaps per year
print(avg_comparisons(mig_lang_diffs, variable = "gender", by = "year", vcov = "HC3"))
```

## Looking more closely at age

### Age and life satisfaction

```{r age_satis}
# Visualize the association between age and satisfaction, by gender and year

# Fully categorical model
# This allows the effect of every year of age to vary freely
# A lot of flexibility but also huge error bars
add_age <- lm(satis ~ gender*as.factor(year)*as.factor(age), 
                  data = combined)
summary(add_age)

# Generate predictions for combinations of gender and year and age
pred <- predictions(add_age,
                 by = c("gender", "year", "age"))

# Visualize
# 2010
cat_age_2010 <- ggplot(pred[pred$year == 2010,], aes(x = age, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("2010") +
  ylab("Mean life satisfaction") +
  xlab("Age") +
  labs(color = "Gender")

# 2015
cat_age_2015 <- ggplot(pred[pred$year == 2015,], aes(x = age, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("2015") +
  ylab("Mean life satisfaction") +
  xlab("Age") +
  labs(color = "Gender")

# 2023
cat_age_2023 <- ggplot(pred[pred$year == 2023,], aes(x = age, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("2023") +
  ylab("Mean life satisfaction") +
  xlab("Age") +
  labs(color = "Gender")

# Combine plots
library(patchwork)
combined_cat_age <- (cat_age_2010|cat_age_2015|cat_age_2023) + plot_layout(guides = "collect")
combined_cat_age
ggsave("Plots/satis_by_age_categorical.png", width = 6, height = 3)


# Re-fit the model including only the linear age effect
# This is the least flexible solution
# But also greatly induces the uncertainty in the estimates
add_age_linear <- lm(satis ~ gender*as.factor(year)*age, 
                  data = combined)
summary(add_age_linear)

# generate predictions for combinations of gender and year and age
pred <- predictions(add_age_linear,
                 by = c("gender", "year", "age"))

# Plot the results
# 2010
lin_age_2010 <- ggplot(pred[pred$year == 2010,], aes(x = age, group = gender, 
                 color = gender,
                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("2010") +
  ylab("Mean life satisfaction") +
  xlab("Age") +
  labs(color = "Gender")

# 2015
lin_age_2015 <- ggplot(pred[pred$year == 2015,], aes(x = age, group = gender, 
                 color = gender,
                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Mean life satisfaction") +
  xlab("Age") +
  labs(color = "Gender") +
  ggtitle("2015")

# 2023
lin_age_2023 <- ggplot(pred[pred$year == 2023,], aes(x = age, group = gender, 
                 color = gender,
                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Mean life satisfaction") +
  xlab("Age") +
  labs(color = "Gender") +
  ggtitle("2023")

# Combine into single plot
combined_lin_age <- (lin_age_2010|lin_age_2015|lin_age_2023) + plot_layout(guides = "collect")
combined_lin_age
ggsave("Plots/satis_by_age_linear.png", width = 6, height = 3)


# Best of both worlds, splines
# Now, we allow for some flexibility
# But we still allow some pooling across years
add_age_splines <- lm(satis ~ gender*as.factor(year)*bs(age, df = 3), 
                  data = combined)
summary(add_age_splines)

# generate predictions for combinations of gender and year and age
pred <- predictions(add_age_splines,
                 by = c("gender", "year", "age"))

# Plot the results
# 2010
spline_age_2010 <- ggplot(pred[pred$year == 2010,], aes(x = age, group = gender, 
                 color = gender,
                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("2010") +
  ylab("Mean life satisfaction") +
  xlab("Age") +
  labs(color = "Gender")

# 2015
spline_age_2015 <- ggplot(pred[pred$year == 2015,], aes(x = age, group = gender, 
                 color = gender,
                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Mean life satisfaction") +
  xlab("Age") +
  labs(color = "Gender") +
  ggtitle("2015")

# 2023
spline_age_2023 <- ggplot(pred[pred$year == 2023,], aes(x = age, group = gender, 
                 color = gender,
                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Mean life satisfaction") +
  xlab("Age") +
  labs(color = "Gender") +
  ggtitle("2023")

# Combine into single plot
combined_spline_age <- (spline_age_2010|spline_age_2015|spline_age_2023) + plot_layout(guides = "collect")
combined_spline_age
ggsave("Plots/satis_by_age_splines.png", width = 6, height = 3)



```

### Does age in isolation explain away the explanandum?

```{r age_explain}
# Fit model including age
# Categorical version here (but in the final model we will use the spline)
explanandum_age <- lm(satis ~ as.factor(year)*gender*as.factor(age), 
                  data = combined)

# New hypothetical data: demographics like in 2010
# This is our counterfactual world in which the sample
# composition did not change with respects to age
# (and any other variable we will include in the model later on)
temp1 <- combined[combined$year == 2010,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2010,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2010,]
temp3$year <- 2023
all2010 <- rbind(temp1, temp2, temp3)


# Generate predictions for combinations of gender and year
pred <- predictions(explanandum_age,
                 by = c("gender", "year"),
                 newdata = all2010)

# Plot the results
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_age_controlled.png", width = 4, height = 3)


# Calculate the gender gaps on the hypothetical data
# In which the demographics remained unchanged across the three survey waves
comps <- avg_comparisons(explanandum_age, variables = "gender", by = "year",
                 newdata = all2010)
print(comps)

# Compare gender gaps 2010 vs 2023
max_comp <- avg_comparisons(explanandum_age, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2010)
print(max_comp)

# Compare gender gaps 2010 vs 2015
print(avg_comparisons(explanandum_age, variables = "gender", by = "year",
                hypothesis = "b1 = b2"),
                 newdata = all2010)

# Compare gender gaps 2015 vs 2023
print(avg_comparisons(explanandum_age, variables = "gender", by = "year",
                hypothesis = "b2 = b3"),
                 newdata = all2010)

# Express differences in SDs
round(comps$estimate[1]/sd(combined$satis), 2)
round(comps$estimate[2]/sd(combined$satis), 2)
round(comps$estimate[3]/sd(combined$satis), 2)

# Express widening gender gap in SDs
round(max_comp$estimate/sd(combined$satis), 2)

# Also calculate widening of gender gap for 2015 and 2023 age distribution

# new data: demographics like in 2015
temp1 <- combined[combined$year == 2015,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2015,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2015,]
temp3$year <- 2023
all2015 <- rbind(temp1, temp2, temp3)
# max comparison for 2015 age distribution
max_comp <- avg_comparisons(explanandum_age, variables = "gender", by = "year",
                hypothesis = "b1 = b3")
print(max_comp)

# new data: demographics like in 2023
temp1 <- combined[combined$year == 2023,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023,]
temp3$year <- 2023
all2023 <- rbind(temp1, temp2, temp3)
# max comparison for 2023 age distribution
max_comp <- avg_comparisons(explanandum_age, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2015)
print(max_comp)
```

### Does age modify the widening of the gender gap?

```{r age_modify}
# Use the spline model
explanandum_age_spline <- lm(satis ~ as.factor(year)*gender*bs(age, df = 3), 
                  data = combined)

# Calculate the gender gaps for each survey year and each age
all_comps <- avg_comparisons(explanandum_age_spline, variables = "gender", by = c("year", "age"))
print(all_comps)

# Calculate the widenings of the gender gap for each age group

# Plot the age trajectory of the widening of the gender gap
widening_per_year <- avg_comparisons(explanandum_age_spline, variables = "gender", by = c("year", "age"),
      hypothesis = c("b15 - b1 = 0", 
                     "b16 - b2 = 0",
                     "b17 - b3 = 0",
                     "b18 - b4 = 0",
                     "b19 - b5 = 0",
                     "b20 - b6 = 0",
                     "b21 - b7 = 0"))
widening_per_year$age <- 12:18

# Plot widening per year
ggplot(widening_per_year, aes(x = age, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  coord_cartesian(ylim = c(-.5, 1)) +
  theme_classic() +
  ylab("Widening of gender gap") +
  xlab("Age")
ggsave("Plots/age_modification.png", width = 4, height = 3)




```

## A closer look at Schooltype

### Schooltype and life satisfaction

```{r schooltype_satis}
# Schooltype as a categorical predictor
add_schooltype <- lm(satis ~ gender*as.factor(year)*as.factor(schooltype), 
                  data = combined)
summary(add_schooltype)

# Generate predictions for combinations of gender and year and age
pred <- predictions(add_schooltype,
                 by = c("gender", "year", "schooltype"))

# Plot the results
# Schooltype 1 (general secondary)
cat_schooltype_1 <- ggplot(pred[pred$schooltype == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("General secondary school") +
  ylab("Mean life satisfaction") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# Schooltype 2 (grammar school)
cat_schooltype_2 <- ggplot(pred[pred$schooltype == 2,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Grammar school") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean life satisfaction") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_schooltype <- (cat_schooltype_1|cat_schooltype_2) + plot_layout(guides = "collect")
combined_cat_schooltype
ggsave("Plots/satis_by_schooltype_categorical.png", width = 6, height = 3)
```

### Does schooltype explain away the explanandum?

```{r schooltype_explain}
# Fit linear model including both age and schooltype
explanandum_schooltype <- lm(satis ~ as.factor(year) + gender + as.factor(year):gender +
                               as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                               as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype), 
                  data = combined)

# Generate predictions for combinations of gender and year
pred <- predictions(explanandum_schooltype,
                 by = c("gender", "year"),
                 newdata = all2010,
                 vcov = ~ unique_classroom)

# Plot the results
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_age_schooltype_controlled.png", width = 4, height = 3)

# Gender gap by year
# For the 2010 schooltype distribution
comps <- avg_comparisons(explanandum_schooltype, variables = "gender", by = "year",
                         newdata = all2010, 
                         vcov = ~ unique_classroom)
print(comps)

# Test widening from 2010 to 2023
max_comp <- avg_comparisons(explanandum_schooltype, variables = "gender", by = "year",
                hypothesis = "b1 = b3", newdata = all2010, vcov = ~ unique_classroom)
print(max_comp)

# Gender gap 2010 vs 2015
print(avg_comparisons(explanandum_schooltype, variables = "gender", by = "year",
                hypothesis = "b1 = b2", newdata = all2010, vcov = ~ unique_classroom))

# Gender gap 2015 vs 2023
print(avg_comparisons(explanandum_schooltype, variables = "gender", by = "year",
                hypothesis = "b2 = b3", newdata = all2010, vcov = ~ unique_classroom))

# Express differences in SDs
round(comps$estimate[1]/sd(combined$satis), 2)
round(comps$estimate[2]/sd(combined$satis), 2)
round(comps$estimate[3]/sd(combined$satis), 2)

# Express widening gender gap in SDs
round(max_comp$estimate/sd(combined$satis), 2)

# Widening of the gap from 2010 to 2023 at 2015 schooltype distribution
max_comp <- avg_comparisons(explanandum_schooltype, variables = "gender", by = "year",
                hypothesis = "b1 = b3", newdata = all2015, vcov = ~ unique_classroom)
print(max_comp)

# Widening of the gap from 2010 to 2023 at 2023 schooltype distribution
max_comp <- avg_comparisons(explanandum_schooltype, variables = "gender", by = "year",
                hypothesis = "b1 = b3", newdata = all2023, vcov = ~ unique_classroom)
print(max_comp)
```

### Does schooltype modify the widening of the gender gap?

```{r schooltype_modify}
# Gender gap for each year and each schooltype
all_comps <- avg_comparisons(explanandum_schooltype, 
                             variables = "gender", 
                             by = c("year", "schooltype"), 
                             vcov = ~ unique_classroom)

# Test widening of gender gap for the lower track students
print(avg_comparisons(explanandum_schooltype, 
                             variables = "gender", 
                             by = c("year", "schooltype"), 
                             vcov = ~ unique_classroom,
                      hypothesis = "b1 = b5")) 

# Test widening of the gender gap for upper track students
print(avg_comparisons(explanandum_schooltype, 
                             variables = "gender", 
                             by = c("year", "schooltype"), 
                             vcov = ~ unique_classroom,
                      hypothesis = "b2 = b6")) # upper track

# Compare widening of the gaps between upper and lower track
# Lower track versus upper track
print(avg_comparisons(explanandum_schooltype, 
                             variables = "gender", 
                             by = c("year", "schooltype"), 
                             vcov = ~ unique_classroom,
                      hypothesis = "b1 - b5 = b2 - b6")) 
```

## Migration background/German not main language

### Migration background and life satisfaction

```{r migback_satis}
# Predict life satisfaction from migration background
add_mig_lang <- lm(satis ~ gender*as.factor(year)*as.factor(mig_lang), 
                  data = combined)
summary(add_mig_lang)

# Generate predictions for combinations of gender and year and age
pred <- predictions(add_mig_lang,
                 by = c("gender", "year", "mig_lang"))

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean life satisfaction") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean life satisfaction") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
ggsave("Plots/satis_by_mig_lang_categorical.png", width = 6, height = 3)
```

### Does migration background explain away the explanandum?

```{r migback_explain}
# Full model including the previous controls
explanandum_mig_lang <- lm(satis ~ as.factor(year) + gender + as.factor(year):gender +
                                   as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                                   as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                   as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang), 
                  data = combined)

# Generate predictions for combinations of gender and year
# Given the 2010 distribution of covariates
pred <- predictions(explanandum_mig_lang,
                 by = c("gender", "year"),
                 newdata = all2010,
                 vcov = ~ unique_classroom)

# Plot the results
ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  labs(color = "Gender")
ggsave("Plots/explanandum_age_schooltype_migrant_controlled.png", width = 4, height = 3)

# Gender gaps by year
comps <- avg_comparisons(explanandum_mig_lang, variables = "gender", by = "year",
                 newdata = all2010,
                 vcov = ~ unique_classroom)
print(comps)

# Test gender gap 2010 vs 2023
max_comp <- avg_comparisons(explanandum_mig_lang, variables = "gender", by = "year",
                hypothesis = "b1 = b3", newdata = all2010, vcov = ~ unique_classroom)
print(max_comp)

# Test gender gap 2010 vs 2015
print(avg_comparisons(explanandum_mig_lang, variables = "gender", by = "year",
                hypothesis = "b1 = b2", newdata = all2010, vcov = ~ unique_classroom))

# Test gender gap 2015 vs 2023
print(avg_comparisons(explanandum_mig_lang, variables = "gender", by = "year",
                hypothesis = "b2 = b3", newdata = all2010, vcov = ~ unique_classroom))

# Express differences in SDs
round(comps$estimate[1]/sd(combined$satis), 2)
round(comps$estimate[2]/sd(combined$satis), 2)
round(comps$estimate[3]/sd(combined$satis), 2)

# Express widening gender gap in SDs
round(max_comp$estimate/sd(combined$satis), 2)

# Compare gender gap 2010 vs 2023, this time for the 2015 distribution of covariates
max_comp <- avg_comparisons(explanandum_mig_lang, variables = "gender", by = "year",
                hypothesis = "b1 = b3", newdata = all2015, vcov = ~ unique_classroom)
print(max_comp)

# Compare gender gap 2010 vs 2023, this time for the 2023 distribution of covariates
max_comp <- avg_comparisons(explanandum_mig_lang, variables = "gender", by = "year",
                hypothesis = "b1 = b3", newdata = all2023, vcov = ~ unique_classroom)
print(max_comp)
```

### Does migration background modify the widening of the gender gap?

```{r migback_modify}
# Calculate gender gaps for each year and each level of migration background
all_comps <- avg_comparisons(explanandum_mig_lang, 
                             variables = "gender", 
                             by = c("year", "mig_lang"), 
                             newdata = all2010,
                             vcov = ~ unique_classroom)
print(all_comps)
# Widening of the gap for no migration background
print(avg_comparisons(explanandum_mig_lang, 
                             variables = "gender", 
                             by = c("year", "mig_lang"), 
                      newdata = all2010,
                             vcov = ~ unique_classroom,
                      hypothesis = "b1 = b5")) 

# Widening of the gap for migration background
print(avg_comparisons(explanandum_mig_lang, 
                             variables = "gender", 
                             by = c("year", "mig_lang"), 
                             vcov = ~ unique_classroom,
                      newdata = all2010,
                      hypothesis = "b2 = b6")) # migback

# Compare the widening of the gap between no migration background and migration background
print(avg_comparisons(explanandum_mig_lang, 
                             variables = "gender", 
                             by = c("year", "mig_lang"), 
                             vcov = ~ unique_classroom,
                      newdata = all2010,
                      hypothesis = "b1 - b5 = b2 - b6")) 

# Widening of the gap for no migration background, 2023 demographics
print(avg_comparisons(explanandum_mig_lang, 
                             variables = "gender", 
                             by = c("year", "mig_lang"), 
                      newdata = all2023,
                             vcov = ~ unique_classroom,
                      hypothesis = "b1 = b5")) 

# Widening of the gap for migration background, 2023 demographics
print(avg_comparisons(explanandum_mig_lang, 
                             variables = "gender", 
                             by = c("year", "mig_lang"), 
                             vcov = ~ unique_classroom,
                      newdata = all2023,
                      hypothesis = "b2 = b6")) # migback

```

## Final model
In the manuscript, we report a final model to address all questions at once.


```{r final}

explanandum_demo_final <- lm(satis ~ as.factor(year) + gender + as.factor(year):gender +
                                   bs(age, df = 3) + as.factor(year):bs(age, df = 3) + gender:bs(age, df = 3) + as.factor(year):gender:bs(age, df = 3) +
                                   as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                   as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang),
                             data = combined)
summary(explanandum_demo_final)
```

### Does the model explain away the explanandum?

```{r final_explain}

# Look at gender gaps for 2010 demographics, by year
comps <- avg_comparisons(explanandum_demo_final, variables = "gender", by = "year",
                         newdata = all2010, 
                         vcov = ~ unique_classroom)
print(comps)

# Test widening from 2010 to 2023
max_comp <- avg_comparisons(explanandum_demo_final, variables = "gender", by = "year",
                hypothesis = "b1 = b3", newdata = all2010, vcov = ~ unique_classroom)
print(max_comp)


# Express widening gender gap in SDs
round(max_comp$estimate/sd(combined$satis), 2)


# Widening of the gap from 2010 to 2023 at 2015 schooltype distribution
max_comp <- avg_comparisons(explanandum_schooltype, variables = "gender", by = "year",
                hypothesis = "b1 = b3", newdata = all2015, vcov = ~ unique_classroom)
print(max_comp)

# Express widening gender gap in SDs
round(max_comp$estimate/sd(combined$satis), 2)

# Widening of the gap from 2010 to 2023 at 2023 schooltype distribution
max_comp <- avg_comparisons(explanandum_schooltype, variables = "gender", by = "year",
                hypothesis = "b1 = b3", newdata = all2023, vcov = ~ unique_classroom)
print(max_comp)

# Express widening gender gap in SDs
round(max_comp$estimate/sd(combined$satis), 2)
```

## Do demographics modify the widening of the gap?

### Modification by age
This will give us Figure 5 from the manuscript.

```{r final_mod_age}

# Calculate margins
all_age_comps <- avg_comparisons(explanandum_demo_final, variables = "gender",
                                 by = c("age", "year"),
                                 vcov = ~ unique_classroom,
                                 newdata = all2010)
print(all_age_comps)

# Calculate widening of the gender gap separately per year of age
widening_per_year <- avg_comparisons(explanandum_age_spline, variables = "gender", by = c("age", "year"), 
      hypothesis = c("b3 - b1 = 0", 
                     "b6 - b4 = 0",
                     "b9 - b7 = 0",
                     "b12 - b10 = 0",
                     "b15 - b13 = 0",
                     "b18 - b16 = 0",
                     "b21 - b19 = 0"),
                      vcov = ~ unique_classroom,
      newdata = all2010)
widening_per_year$age <- 12:18

# Plot widening of the gender gap
ggplot(widening_per_year, aes(x = age, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  coord_cartesian(ylim = c(-.5, 1)) +
  theme_classic() +
  ylab("Widening of gender gap") +
  xlab("Age")
ggsave("Plots/age_modification_final.png", width = 4, height = 3)


# Test maximum contrast
avg_comparisons(explanandum_age_spline, variables = "gender", by = c("age", "year"), 
      hypothesis = "b3 - b1 = b12 - b10",
                      vcov = ~ unique_classroom,
      newdata = all2010)



```

### Modification by schooltype

```{r final_mod_schooltype}
# Calculate margins
all_schooltype_comps <- avg_comparisons(explanandum_demo_final, variables = "gender",
                                                        vcov = ~ unique_classroom,
                                 by = c("schooltype", "year"),
                                 newdata = all2010)
print(all_schooltype_comps)

# Widening among schooltype 1
avg_comparisons(explanandum_demo_final, variables = "gender",
                                 by = c("schooltype", "year"), 
                                vcov = ~ unique_classroom,
                hypothesis = "b1 = b3", newdata = all2010)

# Widening among schooltype 2
avg_comparisons(explanandum_demo_final, variables = "gender",
                                 by = c("schooltype", "year"), 
                                vcov = ~ unique_classroom,
                hypothesis = "b4 = b6", newdata = all2010)


# Interaction
avg_comparisons(explanandum_demo_final, variables = "gender",
                                 by = c("schooltype", "year"), 
                                vcov = ~ unique_classroom,
                hypothesis = "b1 - b3 = b4 - b6", newdata = all2010)

```
### Modification by migration background

```{r final_mod_migback}
# Calculate margins
all_mig_lang_comps <- avg_comparisons(explanandum_demo_final, variables = "gender",
                                      vcov = ~ unique_classroom,
                                      newdata = all2010,
                                 by = c("mig_lang", "year"))
print(all_mig_lang_comps)

# Widening among mig_lang 0
avg_comparisons(explanandum_demo_final, variables = "gender",
                                 by = c("mig_lang", "year"), 
                                vcov = ~ unique_classroom,
                newdata = all2010,
                hypothesis = "b1 = b3")

# Widening among mig_lang 1
avg_comparisons(explanandum_demo_final, variables = "gender",
                                 by = c("mig_lang", "year"), 
                                vcov = ~ unique_classroom,
                newdata = all2010,
                hypothesis = "b4 = b6")


# Interaction
avg_comparisons(explanandum_demo_final, variables = "gender",
                                 by = c("mig_lang", "year"), 
                vcov = ~ unique_classroom,
                newdata = all2010,
                hypothesis = "b1 - b3 = b4 - b6")


```

## Adjusted widening of gender gap, separated by migration background
Generate Figure 6 which summarizes our preliminary results.

```{r plot_adjusted_separated}
# Generate predictions for combinations of gender and year and age
pred <- predictions(explanandum_demo_final,
                    vcov = ~ unique_classroom,
                    newdata = all2010,
                 by = c("gender", "year", "mig_lang"))

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean life satisfaction") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean life satisfaction") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
ggsave("Plots/demo_final_by_mig.png", width = 7, height = 3)


```
