---
title: "Deep dive mode effects"
author: "Julia Rohrer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("prep.RData")
library(haven)
library(marginaleffects)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(patchwork)


combined$source <- as.factor(combined$source)

```

We are going to dive deeper into the mode effects, testing various things that were brought up during a colloquium presentation of the findings.

## What about mode effects on the other items?
We will use the model specification with controls to check for mode effects on the other satisfaction items.
Maybe the mode effects only show up for items that are perceived as somewhat sensitive -- where students may have been reluctant to give honest answers on paper.


```{r mode_satis_controlled}
# satis, for comparison
mode_effects_controlled <- lm(satis ~ source*gender +
                                age + age:gender + age:source +
                                as.factor(schooltype) + as.factor(schooltype):gender + as.factor(schooltype):source +
                                as.factor(mig_lang) + as.factor(mig_lang):gender + as.factor(mig_lang):source, data = combined[combined$year == 2023,])
print(avg_predictions(mode_effects_controlled, variables = c("gender", "source"), vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled, variables = "source", by = "gender", vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled, variables = "source", by = "gender", vcov = ~ unique_classroom,
                hypothesis = "b1 = b2"))
# satis_money
# matches the pattern
mode_effects_controlled_money <- lm(satis_money ~ source*gender +
                                age + age:gender + age:source +
                                as.factor(schooltype) + as.factor(schooltype):gender + as.factor(schooltype):source +
                                as.factor(mig_lang) + as.factor(mig_lang):gender + as.factor(mig_lang):source, data = combined[combined$year == 2023,])
print(avg_predictions(mode_effects_controlled_money, variables = c("gender", "source"), vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_money, variables = "source", by = "gender", vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_money, variables = "source", by = "gender", vcov = ~ unique_classroom,
                hypothesis = "b1 = b2"))

# satis_friends
mode_effects_controlled_friends <- lm(satis_friends ~ source*gender +
                                age + age:gender + age:source +
                                as.factor(schooltype) + as.factor(schooltype):gender + as.factor(schooltype):source +
                                as.factor(mig_lang) + as.factor(mig_lang):gender + as.factor(mig_lang):source, data = combined[combined$year == 2023,])
print(avg_predictions(mode_effects_controlled_friends, variables = c("gender", "source"), vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_friends, variables = "source", by = "gender", vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_friends, variables = "source", by = "gender", vcov = ~ unique_classroom,
                hypothesis = "b1 = b2"))

# satis_mom
mode_effects_controlled_mom <- lm(satis_mom ~ source*gender +
                                age + age:gender + age:source +
                                as.factor(schooltype) + as.factor(schooltype):gender + as.factor(schooltype):source +
                                as.factor(mig_lang) + as.factor(mig_lang):gender + as.factor(mig_lang):source, data = combined[combined$year == 2023,])
print(avg_predictions(mode_effects_controlled_mom, variables = c("gender", "source"), vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_mom, variables = "source", by = "gender", vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_mom, variables = "source", by = "gender", vcov = ~ unique_classroom,
                hypothesis = "b1 = b2"))


# satis_dad
mode_effects_controlled_dad <- lm(satis_dad ~ source*gender +
                                age + age:gender + age:source +
                                as.factor(schooltype) + as.factor(schooltype):gender + as.factor(schooltype):source +
                                as.factor(mig_lang) + as.factor(mig_lang):gender + as.factor(mig_lang):source, data = combined[combined$year == 2023,])
print(avg_predictions(mode_effects_controlled_dad, variables = c("gender", "source"), vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_dad, variables = "source", by = "gender", vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_dad, variables = "source", by = "gender", vcov = ~ unique_classroom,
                hypothesis = "b1 = b2"))

# satis_leisure
# Boys happier with their leisure time activities when responding on tablet
mode_effects_controlled_leisure <- lm(satis_leisure ~ source*gender +
                                age + age:gender + age:source +
                                as.factor(schooltype) + as.factor(schooltype):gender + as.factor(schooltype):source +
                                as.factor(mig_lang) + as.factor(mig_lang):gender + as.factor(mig_lang):source, data = combined[combined$year == 2023,])
print(avg_predictions(mode_effects_controlled_leisure, variables = c("gender", "source"), vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_leisure, variables = "source", by = "gender", vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_leisure, variables = "source", by = "gender", vcov = ~ unique_classroom,
                hypothesis = "b1 = b2"))

# satis_dwell
mode_effects_controlled_dwell <- lm(satis_dwell ~ source*gender +
                                age + age:gender + age:source +
                                as.factor(schooltype) + as.factor(schooltype):gender + as.factor(schooltype):source +
                                as.factor(mig_lang) + as.factor(mig_lang):gender + as.factor(mig_lang):source, data = combined[combined$year == 2023,])
print(avg_predictions(mode_effects_controlled_dwell, variables = c("gender", "source"), vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_dwell, variables = "source", by = "gender", vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_dwell, variables = "source", by = "gender", vcov = ~ unique_classroom,
                hypothesis = "b1 = b2"))

# satis_grades
# cross over interaction
mode_effects_controlled_grades <- lm(satis_grades ~ source*gender +
                                age + age:gender + age:source +
                                as.factor(schooltype) + as.factor(schooltype):gender + as.factor(schooltype):source +
                                as.factor(mig_lang) + as.factor(mig_lang):gender + as.factor(mig_lang):source, data = combined[combined$year == 2023,])
print(avg_predictions(mode_effects_controlled_grades, variables = c("gender", "source"), vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_grades, variables = "source", by = "gender", vcov = ~ unique_classroom))
print(avg_comparisons(mode_effects_controlled_grades, variables = "source", by = "gender", vcov = ~ unique_classroom,
                hypothesis = "b1 = b2"))

```

## Can the mode effect be explained by differential media usage -- gaming versus social media?


## Is the mode effect age specific which could point toward issues of familiarity?


## Does life satisfaction vary by survey mode?
First, we will simply see whether gender and mode jointly predict life satisfaction in 2023.
```{r mode_satis}
mode_effects <- lm(satis ~ source*gender, data = combined[combined$year == 2023,])
summary(mode_effects)

pred <- avg_predictions(mode_effects, variables = c("source", "gender"), vcov = ~ unique_classroom)

ggplot(pred, aes(x = source, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .2) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey mode") +
  labs(color = "Gender")
ggsave("Plots/mode.png", width = 4, height = 3)


print(avg_predictions(mode_effects, variables = c("source", "gender"),
                hypothesis = "b1 = b3"), vcov = ~ unique_classroom) # Girls: Paper versus online

print(avg_predictions(mode_effects, variables = c("source", "gender"),
                hypothesis = "b2 = b4"), vcov = ~ unique_classroom) # Boys: paper versus online

# Girls are less satisfied on the tablet
# Boys are pretty much the same

# Compare the predictions
print(avg_predictions(mode_effects, variables = c("source", "gender", "year"),
                hypothesis = "b1-b3 = b2-b4", vcov = ~ unique_classroom))

```

## Let's look at whether survey mode covaries with other variables
Maybe survey mode predicts life satisfaction because it covaries with other factors?
To check this possibility, let's see how the samples differ

```{r distribution}
# Extract 2023 final data for convenience
combined_2023 <- combined[combined$year == 2023,]
prop.table(table(combined_2023$source))

# Which schools have which assessment mode(s)
has_tablet <- unique(combined_2023$school_id[combined_2023$source == "online" & !is.na(combined_2023$school_id)])
has_paper <- unique(combined_2023$school_id[combined_2023$source == "paper" & !is.na(combined_2023$school_id)])
has_tablet
has_paper

# Assessment mode table
assessment_modes <- data.frame(matrix(NA, ncol = 3,nrow = length(unique(combined_2023$school_id[!is.na(combined_2023$school_id)]))))
names(assessment_modes) <- c("school_id", "has_paper", "has_tablet")
assessment_modes$school_id <- unique(combined_2023$school_id[!is.na(combined_2023$school_id)])

for (i in 1:nrow(assessment_modes)) {
  assessment_modes$has_paper[i] <-  assessment_modes$school_id[i] %in% has_paper
  assessment_modes$has_tablet[i] <-  assessment_modes$school_id[i] %in% has_tablet
}

assessment_modes$count <- assessment_modes$has_paper + assessment_modes$has_tablet
prop.table(table(assessment_modes$count))
# 63 percent of schools had only one mode, 36 percent of schools had both

# Is response mode correlated with the type of school?
prop.table(table(combined_2023$schooltype, combined_2023$source), margin = 1)
chisq.test(table(combined_2023$schooltype, combined_2023$source))
# Significantly more tablets in Mittelschule

# Is response mode correlated with age?
prop.table(table(combined_2023$age, combined_2023$source), margin = 1)
t.test(combined_2023$age[combined_2023$source == "paper"], combined_2023$age[combined_2023$source == "online"])
mean(combined_2023$age[combined_2023$source == "paper"], na.rm = TRUE)
mean(combined_2023$age[combined_2023$source == "online"], na.rm = TRUE)
# Students who filled out on tablet are slightly but significantly older

# Is response mode correlated with migration background?
prop.table(table(combined_2023$mig_lang, combined_2023$source), margin = 1)
chisq.test(table(combined_2023$mig_lang, combined_2023$source))

```


## School fixed effects
Of course, survey mode may also correlate with some other unobserved variable (e.g., SES) that ultimately explains its associations with life satisfaction.
As survey mode was randomized, we cannot rule out this possibility.
However, let's assume that within schools, these unobserved variables are approximately constant.
In that scenario, we can compare classrooms that used different modes within the same school.

So, if we still find that survey mode is associated with life satisfaction, we can rule out that the association is explained away by unobserved variables on the school level.

```{r fixed_effects}
# Determine which schools used both assessment modes
schools_with_both <- assessment_modes$school_id[assessment_modes$count == 2]

# Run a fixed effects model
# These model include a school-wise intercept 
# and allow the effect of gender to vary between schools
# (in completely arbitrary ways)

mode_effects_controlled_fe <- lm(satis ~ source*gender +
                                   age + age:gender + age:source +
                                   as.factor(schooltype) + as.factor(schooltype):gender + as.factor(schooltype):source +
                                   as.factor(mig_lang) + as.factor(mig_lang):gender + as.factor(mig_lang):source +
                                   as.factor(school_id) + as.factor(school_id):gender, 
                                 data = combined[combined$year == 2023 & combined$school_id %in% schools_with_both,])
summary(mode_effects_controlled_fe)

pred <- avg_predictions(mode_effects_controlled_fe, variables = c("source", "gender"), vcov = ~ unique_classroom)

ggplot(pred, aes(x = source, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .2) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey mode") +
  labs(color = "Gender")
ggsave("Plots/mode_fe.png", width = 4, height = 3)

# Let's again look at mode effects by gender
contrasts <- avg_comparisons(mode_effects_controlled_fe, variables = "source", by = "gender", vcov = ~ unique_classroom)
print(contrasts)

# Once again, girls are less satisfied on tablet
# Boys are pretty much the same

print(avg_comparisons(mode_effects_controlled_fe, variables = "source", by = "gender", vcov = ~ unique_classroom,
                hypothesis = "b1 = b2"))
# Difference of 0.25, p < .03

# This speaks against the idea that survey mode is correlated with some unobserved school-level
# variable that ultimately explains differences in the gender gaps
```



## What happens to the explanandum if we take into account survey mode?
So, let's assume that the (gender-specific) associations between survey mode and life satisfaction actually reflect mode effects: answering the survey on a tablet makes girls (but not boys) report lower levels of satisfaction.
What does the widening of the gender gap look like if we hold survey mode constant?
Tablets were only introduced in 2023, and so we cannot see how the gender gap widens for this assessment mode.
However, paper was used throughout, and so we can look at the widening of the gender gap for this survey mode

```{r mode_explanandum}
explanandum_paper_only <- lm(satis ~ as.factor(year) + gender + as.factor(year):gender +
                                 as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                                 as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                 as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang), 
                  data = combined[combined$source == "paper",])

# We will evaluate this for various data grids
# Demographics like in 2010
temp1 <- combined[combined$year == 2010,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2010,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2010,]
temp3$year <- 2023
all2010 <- rbind(temp1, temp2, temp3)

# Demographics like in 2015
temp1 <- combined[combined$year == 2015,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2015,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2015,]
temp3$year <- 2023
all2015 <- rbind(temp1, temp2, temp3)

# Demographics like in 2023
temp1 <- combined[combined$year == 2023,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023,]
temp3$year <- 2023
all2023 <- rbind(temp1, temp2, temp3)

# Demographics like 2023, but only those who actually used paper
temp1 <- combined[combined$year == 2023 & combined$source == "paper",]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023 & combined$source == "paper",]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023 & combined$source == "paper",]
temp3$year <- 2023
all2023_paper <- rbind(temp1, temp2, temp3)


##########################
# Use 2010 distribution
##########################

max_comp <- avg_comparisons(explanandum_paper_only, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2010,
                vcov = ~ unique_classroom)
print(max_comp)
# Gap from previously 0.27 to 0.17

##########################
# Use 2015 distribution
##########################
max_comp <- avg_comparisons(explanandum_paper_only, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2015,
                vcov = ~ unique_classroom)
print(max_comp)
# Gap from 0.33 to 0.24

##########################
# Use 2023 distribution
##########################
pred <- predictions(explanandum_paper_only,
                 by = c("gender", "year"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)


comps <- avg_comparisons(explanandum_paper_only, variables = "gender", by = "year",
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps)

max_comp <- avg_comparisons(explanandum_paper_only, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2023,
                vcov = ~ unique_classroom)
print(max_comp)
# Gap from 0.31 to 0.21
```
## Paper only, modification by migback
Before, we had already noted that the widening of the gender gap strongly differed depending on migration background/language at home.
So let's see how migration background modifies the widening of the gender gap if we hold constant survey mode by only analyzing responses collected on paper.

```{r mode_migback_modify}
comps_by_migback <- avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2010,
                 vcov = ~ unique_classroom)
print(comps_by_migback)

# Widening of the gap for no migration background
# 2010
print(avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2010,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))
# 0.1, p = .363

# 2015
print(avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2015,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))
# 0.17, p = .132

# 2023
print(avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))
# 0.12, p = .25

print(avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023_paper,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))
# 0.09, p = .384

# Widening of the gap for migration background
# 2010
print(avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2010,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))
# .87, p = .004

# Comparison of the widening
print(avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 - b5 = b2 - b6"))


# 2015
print(avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2015,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))
# 0.93, .003

# 2023
print(avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))
# 0.88, .004
print(avg_comparisons(explanandum_paper_only, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023_paper,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))
# 0.87, p = .004

```
## Plot the implied means of this final model
```{r final_plot}
# Generate predictions for combinations of gender and year and age
pred <- avg_predictions(explanandum_paper_only,
                        newdata = all2023,
                        by = c("gender", "year", "mig_lang"),
                        vcov = ~ unique_classroom)

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean life satisfaction") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean life satisfaction") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
ggsave("Plots/paper_satis_by_mig_lang_categorical.png", width = 6, height = 3)


```

