---
title: "Look at pattern in various groups"
author: "Julia Rohrer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("prep.RData")
library(haven)
library(marginaleffects)
library(ggplot2)
library(ordinal)
library(gridExtra)
```

## Agenda
We are now going to contrast the pattern across various subgroups to get a better feel for what is going on.
We will split analyses by:
* Migration background
* Schooltype

And report:
* statistical significance of the interaction
* the means plot in raw scores
* and an effect size measure based on the overall standardizes satisfaction score
* results from an ordinal model

Furthermore, we will consider:
* all relevant outcomes


## For reference, the whole sample

```{r }

simple_model <- lm(satis ~ as.factor(year)*gender, data = combined, weights = wph)
# Unstandardized estimate
round(summary(simple_model)$coefficients["as.factor(year)2023:gendermale",], 3)
# Estimate standardized relative to total SD
round(summary(simple_model)$coefficients["as.factor(year)2023:gendermale", "Estimate"]/sd(combined$satis, na.rm = TRUE), 3)

# Corresponding ordinal results
ordinal <- clm(as.factor(satis) ~ as.factor(year)*gender, data = combined, weights = wph)
round(summary(ordinal)$coefficients["as.factor(year)2023:gendermale",], 3)


pred <- predictions(simple_model,
                 by = c("gender", "year"))

ggplot(pred, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year")



```

## Let's split by school type

```{r }

simple_model_schooltype1 <- lm(satis ~ as.factor(year)*gender, data = combined[combined$schooltype == 1,], weights = wph)
simple_model_schooltype2 <- lm(satis ~ as.factor(year)*gender, data = combined[combined$schooltype == 2,], weights = wph)

# Unstandardized estimate
# Schooltype == 1
round(summary(simple_model_schooltype1)$coefficients["as.factor(year)2023:gendermale",], 3)
# Schooltype == 2
round(summary(simple_model_schooltype2)$coefficients["as.factor(year)2023:gendermale",], 3)

# Estimate standardized relative to total SD
# Schooltype == 1
round(summary(simple_model_schooltype1)$coefficients["as.factor(year)2023:gendermale", "Estimate"]/sd(combined$satis, na.rm = TRUE), 3)
# Schooltype == 2
round(summary(simple_model_schooltype2)$coefficients["as.factor(year)2023:gendermale", "Estimate"]/sd(combined$satis, na.rm = TRUE), 3)

# Corresponding ordinal results
# Schooltype == 1
ordinal1 <- clm(as.factor(satis) ~ as.factor(year)*gender, data = combined[combined$schooltype == 1,], weights = wph)
round(summary(ordinal1)$coefficients["as.factor(year)2023:gendermale",], 3)
# Schooltype == 2
ordinal2 <- clm(as.factor(satis) ~ as.factor(year)*gender, data = combined[combined$schooltype == 2,], weights = wph)
round(summary(ordinal2)$coefficients["as.factor(year)2023:gendermale",], 3)

pred1 <- predictions(simple_model_schooltype1,
                 by = c("gender", "year"))
pred2 <- predictions(simple_model_schooltype2,
                 by = c("gender", "year"))

plot1 <- ggplot(pred1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  ggtitle("Schooltype Mittelschule")

plot2 <- ggplot(pred2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  ggtitle("Schooltype Gymnasium")

grid.arrange(plot1, plot2, ncol = 2)



```

## Let's split by languate at home

```{r}
# German at home == 1
simple_model_german1 <- lm(satis ~ as.factor(year)*gender, data = combined[combined$german_at_home == 1,], weights = wph)
# German at home == 0
simple_model_german0 <- lm(satis ~ as.factor(year)*gender, data = combined[combined$german_at_home == 0,], weights = wph)

# Unstandardized estimate
# German at home == 1
round(summary(simple_model_german1)$coefficients["as.factor(year)2023:gendermale",], 3)
# German at home == 0
round(summary(simple_model_german0)$coefficients["as.factor(year)2023:gendermale",], 3)

# Estimate standardized relative to total SD
# German at home == 1
round(summary(simple_model_german1)$coefficients["as.factor(year)2023:gendermale", "Estimate"]/sd(combined$satis, na.rm = TRUE), 3)
# German at home == 0
round(summary(simple_model_german0)$coefficients["as.factor(year)2023:gendermale", "Estimate"]/sd(combined$satis, na.rm = TRUE), 3)

# Corresponding ordinal results
# German at home == 1
ordinal1 <- clm(as.factor(satis) ~ as.factor(year)*gender, data = combined[combined$german_at_home == 1,], weights = wph)
round(summary(ordinal1)$coefficients["as.factor(year)2023:gendermale",], 3)
# German at home == 0
ordinal2 <- clm(as.factor(satis) ~ as.factor(year)*gender, data = combined[combined$german_at_home == 0,], weights = wph)
round(summary(ordinal2)$coefficients["as.factor(year)2023:gendermale",], 3)

pred1 <- predictions(simple_model_german1,
                 by = c("gender", "year"))
pred2 <- predictions(simple_model_german0,
                 by = c("gender", "year"))

plot1 <- ggplot(pred1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  ggtitle("German at home")

plot2 <- ggplot(pred2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  ggtitle("No German at home")

grid.arrange(plot1, plot2, ncol = 2)



```

## Let's split by more detailed migration background

```{r }
# Three groups:
# None
# both/self
# father or mother

# migback none
migback_none <- lm(satis ~ as.factor(year)*gender, data = combined[combined$migback == "none",], weights = wph)
# migback both
migback_both <- lm(satis ~ as.factor(year)*gender, data = combined[combined$migback == "both" |combined$migback == "self",], weights = wph)
# migback one
migback_some <- lm(satis ~ as.factor(year)*gender, data = combined[combined$migback == "father" |combined$migback == "mother",], weights = wph)


# Unstandardized estimate
# migback none
round(summary(migback_none)$coefficients["as.factor(year)2023:gendermale",], 3)
# migback both
round(summary(migback_both)$coefficients["as.factor(year)2023:gendermale",], 3)
# migback one
round(summary(migback_some)$coefficients["as.factor(year)2023:gendermale",], 3)

# Estimate standardized relative to total SD
# migback none
round(summary(migback_none)$coefficients["as.factor(year)2023:gendermale", "Estimate"]/sd(combined$satis, na.rm = TRUE), 3)
# migback both
round(summary(migback_both)$coefficients["as.factor(year)2023:gendermale", "Estimate"]/sd(combined$satis, na.rm = TRUE), 3)
# migback one
round(summary(migback_some)$coefficients["as.factor(year)2023:gendermale", "Estimate"]/sd(combined$satis, na.rm = TRUE), 3)

# Corresponding ordinal results
# migback none
ordinal1 <- clm(as.factor(satis) ~ as.factor(year)*gender, data = combined[combined$migback == "none",], weights = wph)
round(summary(ordinal1)$coefficients["as.factor(year)2023:gendermale",], 3)
# migback both
ordinal2 <- clm(as.factor(satis) ~ as.factor(year)*gender, data = combined[combined$migback == "both" |combined$migback == "self",], weights = wph)
round(summary(ordinal2)$coefficients["as.factor(year)2023:gendermale",], 3)
# migback one
ordinal3 <- clm(as.factor(satis) ~ as.factor(year)*gender, data = combined[combined$migback == "father" |combined$migback == "mother",], weights = wph)
round(summary(ordinal3)$coefficients["as.factor(year)2023:gendermale",], 3)

pred1 <- predictions(migback_none,
                 by = c("gender", "year"))
pred2 <- predictions(migback_both,
                 by = c("gender", "year"))
pred3 <- predictions(migback_some,
                 by = c("gender", "year"))

plot1 <- ggplot(pred1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  ggtitle("No migback")

plot2 <- ggplot(pred2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  ggtitle("Full migback")

plot3 <- ggplot(pred3, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean life satisfaction") +
  xlab("Survey year") +
  ggtitle("Half migback")

grid.arrange(plot1, plot2, plot3, ncol = 3)



```

## Let's do the other outcomes by language at home


```{r, fig.width = 10, fig.height = 40}

satis_money_german1 <- lm(satis_money ~ as.factor(year)*gender, data = combined[combined$german_at_home == 1,], weights = wph)
satis_money_german0 <- lm(satis_money ~ as.factor(year)*gender, data = combined[combined$german_at_home == 0,], weights = wph)
pred_money_german1 <- predictions(satis_money_german1, by = c("gender", "year"))
pred_money_german0 <- predictions(satis_money_german0, by = c("gender", "year"))

satis_friends_german1 <- lm(satis_friends ~ as.factor(year)*gender, data = combined[combined$german_at_home == 1,], weights = wph)
satis_friends_german0 <- lm(satis_friends ~ as.factor(year)*gender, data = combined[combined$german_at_home == 0,], weights = wph)
pred_friends_german1 <- predictions(satis_friends_german1, by = c("gender", "year"))
pred_friends_german0 <- predictions(satis_friends_german0, by = c("gender", "year"))

satis_mom_german1 <- lm(satis_mom ~ as.factor(year)*gender, data = combined[combined$german_at_home == 1,], weights = wph)
satis_mom_german0 <- lm(satis_mom ~ as.factor(year)*gender, data = combined[combined$german_at_home == 0,], weights = wph)
pred_mom_german1 <- predictions(satis_mom_german1, by = c("gender", "year"))
pred_mom_german0 <- predictions(satis_mom_german0, by = c("gender", "year"))

satis_dad_german1 <- lm(satis_dad ~ as.factor(year)*gender, data = combined[combined$german_at_home == 1,], weights = wph)
satis_dad_german0 <- lm(satis_dad ~ as.factor(year)*gender, data = combined[combined$german_at_home == 0,], weights = wph)
pred_dad_german1 <- predictions(satis_dad_german1, by = c("gender", "year"))
pred_dad_german0 <- predictions(satis_dad_german0, by = c("gender", "year"))

satis_leisure_german1 <- lm(satis_leisure ~ as.factor(year)*gender, data = combined[combined$german_at_home == 1,], weights = wph)
satis_leisure_german0 <- lm(satis_leisure ~ as.factor(year)*gender, data = combined[combined$german_at_home == 0,], weights = wph)
pred_leisure_german1 <- predictions(satis_leisure_german1, by = c("gender", "year"))
pred_leisure_german0 <- predictions(satis_leisure_german0, by = c("gender", "year"))

satis_dwell_german1 <- lm(satis_dwell ~ as.factor(year)*gender, data = combined[combined$german_at_home == 1,], weights = wph)
satis_dwell_german0 <- lm(satis_dwell ~ as.factor(year)*gender, data = combined[combined$german_at_home == 0,], weights = wph)
pred_dwell_german1 <- predictions(satis_dwell_german1, by = c("gender", "year"))
pred_dwell_german0 <- predictions(satis_dwell_german0, by = c("gender", "year"))

satis_grades_german1 <- lm(satis_grades ~ as.factor(year)*gender, data = combined[combined$german_at_home == 1,], weights = wph)
satis_grades_german0 <- lm(satis_grades ~ as.factor(year)*gender, data = combined[combined$german_at_home == 0,], weights = wph)
pred_grades_german1 <- predictions(satis_grades_german1, by = c("gender", "year"))
pred_grades_german0 <- predictions(satis_grades_german0, by = c("gender", "year"))



money_1 <- ggplot(pred_money_german1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("Money, German at home")

money_2 <- ggplot(pred_money_german0, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("Money, No german at home")

friends_1 <- ggplot(pred_friends_german1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("friends, German at home")

friends_2 <- ggplot(pred_friends_german0, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("friends, No German at home")

mom_1 <- ggplot(pred_mom_german1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("mom, German at home")

mom_2 <- ggplot(pred_mom_german0, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("mom, No German at home")

dad_1 <- ggplot(pred_dad_german1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("dad, German at home")

dad_2 <- ggplot(pred_dad_german0, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("dad, No German at home")

leisure_1 <- ggplot(pred_leisure_german1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("leisure, German at home")

leisure_2 <- ggplot(pred_leisure_german0, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("leisure, No German at home")

dwell_1 <- ggplot(pred_dwell_german1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("dwell, German at home")

dwell_2 <- ggplot(pred_dwell_german0, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("dwell, No German at home")

grades_1 <- ggplot(pred_grades_german1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("grades, German at home")

grades_2 <- ggplot(pred_grades_german0, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("grades, No German at home")

grid.arrange(money_1, money_2, 
             friends_1, friends_2,
             mom_1, mom_2,
             dad_1, dad_2,
             leisure_1, leisure_2,
             dwell_1, dwell_2,
             grades_1, grades_2, ncol = 2)



```

## Let's do the other outcomes by school type
Given the results for general satisfaction, I do not really expect much to be going on here.


```{r, fig.width = 10, fig.height = 40}

satis_money_schooltype1 <- lm(satis_money ~ as.factor(year)*gender, data = combined[combined$schooltype == 1,], weights = wph)
satis_money_schooltype2 <- lm(satis_money ~ as.factor(year)*gender, data = combined[combined$schooltype == 2,], weights = wph)
pred_money_schooltype1 <- predictions(satis_money_schooltype1, by = c("gender", "year"))
pred_money_schooltype2 <- predictions(satis_money_schooltype2, by = c("gender", "year"))

satis_friends_schooltype1 <- lm(satis_friends ~ as.factor(year)*gender, data = combined[combined$schooltype == 1,], weights = wph)
satis_friends_schooltype2 <- lm(satis_friends ~ as.factor(year)*gender, data = combined[combined$schooltype == 2,], weights = wph)
pred_friends_schooltype1 <- predictions(satis_friends_schooltype1, by = c("gender", "year"))
pred_friends_schooltype2 <- predictions(satis_friends_schooltype2, by = c("gender", "year"))

satis_mom_schooltype1 <- lm(satis_mom ~ as.factor(year)*gender, data = combined[combined$schooltype == 1,], weights = wph)
satis_mom_schooltype2 <- lm(satis_mom ~ as.factor(year)*gender, data = combined[combined$schooltype == 2,], weights = wph)
pred_mom_schooltype1 <- predictions(satis_mom_schooltype1, by = c("gender", "year"))
pred_mom_schooltype2 <- predictions(satis_mom_schooltype2, by = c("gender", "year"))

satis_dad_schooltype1 <- lm(satis_dad ~ as.factor(year)*gender, data = combined[combined$schooltype == 1,], weights = wph)
satis_dad_schooltype2 <- lm(satis_dad ~ as.factor(year)*gender, data = combined[combined$schooltype == 2,], weights = wph)
pred_dad_schooltype1 <- predictions(satis_dad_schooltype1, by = c("gender", "year"))
pred_dad_schooltype2 <- predictions(satis_dad_schooltype2, by = c("gender", "year"))

satis_leisure_schooltype1 <- lm(satis_leisure ~ as.factor(year)*gender, data = combined[combined$schooltype == 1,], weights = wph)
satis_leisure_schooltype2 <- lm(satis_leisure ~ as.factor(year)*gender, data = combined[combined$schooltype == 2,], weights = wph)
pred_leisure_schooltype1 <- predictions(satis_leisure_schooltype1, by = c("gender", "year"))
pred_leisure_schooltype2 <- predictions(satis_leisure_schooltype2, by = c("gender", "year"))

satis_dwell_schooltype1 <- lm(satis_dwell ~ as.factor(year)*gender, data = combined[combined$schooltype == 1,], weights = wph)
satis_dwell_schooltype2 <- lm(satis_dwell ~ as.factor(year)*gender, data = combined[combined$schooltype == 2,], weights = wph)
pred_dwell_schooltype1 <- predictions(satis_dwell_schooltype1, by = c("gender", "year"))
pred_dwell_schooltype2 <- predictions(satis_dwell_schooltype2, by = c("gender", "year"))

satis_grades_schooltype1 <- lm(satis_grades ~ as.factor(year)*gender, data = combined[combined$schooltype == 1,], weights = wph)
satis_grades_schooltype2 <- lm(satis_grades ~ as.factor(year)*gender, data = combined[combined$schooltype == 2,], weights = wph)
pred_grades_schooltype1 <- predictions(satis_grades_schooltype1, by = c("gender", "year"))
pred_grades_schooltype2 <- predictions(satis_grades_schooltype2, by = c("gender", "year"))



money_1 <- ggplot(pred_money_schooltype1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("Money, Schooltype Mittelschule")

money_2 <- ggplot(pred_money_schooltype2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("Money, Schooltype Gymnasium")

friends_1 <- ggplot(pred_friends_schooltype1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("friends, Schooltype Mittelschule")

friends_2 <- ggplot(pred_friends_schooltype2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("friends, Schooltype Gymnasium")

mom_1 <- ggplot(pred_mom_schooltype1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("mom, Schooltype Mittelschule")

mom_2 <- ggplot(pred_mom_schooltype2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("mom, Schooltype Gymnasium")

dad_1 <- ggplot(pred_dad_schooltype1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("dad, Schooltype Mittelschule")

dad_2 <- ggplot(pred_dad_schooltype2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("dad, Schooltype Gymnasium")

leisure_1 <- ggplot(pred_leisure_schooltype1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("leisure, Schooltype Mittelschule")

leisure_2 <- ggplot(pred_leisure_schooltype2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("leisure, Schooltype Gymnasium")

dwell_1 <- ggplot(pred_dwell_schooltype1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("dwell, Schooltype Mittelschule")

dwell_2 <- ggplot(pred_dwell_schooltype2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("dwell, Schooltype Gymnasium")

grades_1 <- ggplot(pred_grades_schooltype1, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("grades, Schooltype Mittelschule")

grades_2 <- ggplot(pred_grades_schooltype2, aes(x = year, group = gender, color = gender, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = .5) +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  coord_cartesian(ylim = c(2.5, 5)) +
  theme_classic() +
  ylab("Mean satisfaction") +
  xlab("Survey year") +
  ggtitle("grades, Schooltype Gymnasium")

grid.arrange(money_1, money_2, 
             friends_1, friends_2,
             mom_1, mom_2,
             dad_1, dad_2,
             leisure_1, leisure_2,
             dwell_1, dwell_2,
             grades_1, grades_2, ncol = 2)



```
