---
title: "Mode effects and school fixed effects"
author: "Julia Rohrer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("prep.RData")
library(haven)
library(marginaleffects)
library(ggplot2)
library(gridExtra)
library(dplyr)

```

## Agenda
In 2023, the survey was conducted with tablets whenever they were available in the schools.
Here, we are going to look at the possibility that this might have affected how students report their life satisfaction.


## Let's look at the distribution of response mode in 2023

```{r }
# Extract 2023 final data for convenience
combined_2023 <- combined[combined$year == 2023,]
prop.table(table(combined_2023$source))

# Schools with assessment modes
has_tablet <- unique(combined_2023$school_id[combined_2023$source == "online" & !is.na(combined_2023$school_id)])
has_paper <- unique(combined_2023$school_id[combined_2023$source == "paper" & !is.na(combined_2023$school_id)])
has_tablet
has_paper

# Assessment mode table
assessment_modes <- data.frame(matrix(NA, ncol = 3,nrow = length(unique(combined_2023$school_id[!is.na(combined_2023$school_id)]))))
names(assessment_modes) <- c("school_id", "has_paper", "has_tablet")
assessment_modes$school_id <- unique(combined_2023$school_id[!is.na(combined_2023$school_id)])

for (i in 1:nrow(assessment_modes)) {
  assessment_modes$has_paper[i] <-  assessment_modes$school_id[i] %in% has_paper
  assessment_modes$has_tablet[i] <-  assessment_modes$school_id[i] %in% has_tablet
}

assessment_modes$count <- assessment_modes$has_paper + assessment_modes$has_tablet
prop.table(table(assessment_modes$count))
# 63 percent of schools had only one mode, 36 percent of schools had both


# Is response mode correlated with the type of school?
prop.table(table(combined_2023$schooltype, combined_2023$source), margin = 1)
# Many more tablets at Mittelschule

# Is response mode correlated with the age (proxy for grade level)?
prop.table(table(combined_2023$age, combined_2023$source), margin = 1)
t.test(combined_2023$age[combined_2023$source == "paper"], combined_2023$age[combined_2023$source == "online"])
# Highly significant association
mean(combined_2023$age[combined_2023$source == "paper"], na.rm = TRUE)
mean(combined_2023$age[combined_2023$source == "online"], na.rm = TRUE)
# Students who filled out on tablet are slightly older
```
## Does life satisfaction vary by response mode?
```{r }
mode_effects <- lm(satis ~ as.factor(source)*gender, data = combined_2023)
summary(mode_effects)
# there is a weak interaction here
pred_mode_effects <- predictions(mode_effects,
                 by = c("gender", "source"))
pred_mode_effects
# Girls are less satisfied on the tabled
# Boys are pretty much the same

# What if we include school fixed effects?
mode_effects_school <- lm(satis ~ as.factor(source)*gender + as.factor(school_id), data = combined[combined$year == 2023,])
summary(mode_effects_school)
# there is a weak interaction here
pred_mode_effects_school <- predictions(mode_effects_school,
                 by = c("gender", "source"))
pred_mode_effects_school

# What about if we allow the gender-gap to vary by school?
mode_effects_school <- lm(satis ~ as.factor(source)*gender + gender*as.factor(school_id), data = combined_2023)
summary(mode_effects_school)
pred_mode_effects_school <- predictions(mode_effects_school,
                 by = c("gender", "source"))
pred_mode_effects_school
# there is still weak interaction here

# Let's additionally account age (schooltype is already taken care of by school fixed effect)
mode_effects_school_plus <- lm(satis ~ as.factor(source)*gender + gender*as.factor(school_id) + gender*age, data = combined_2023)
summary(mode_effects_school_plus)
# there is still an interaction here


```


## Do we get the same patterns for the other outcomes?
```{r }
# General satisfaction
mode_effects_school_plus <- lm(satis ~ as.factor(source)*gender + gender*as.factor(school_id) + gender*age, data = combined_2023)
summary(mode_effects_school_plus)$coefficients["as.factor(source)paper:gendermale",]

# Satisfaction with money
mode_effects_school_plus <- lm(satis_money ~ as.factor(source)*gender + gender*as.factor(school_id) + gender*age, data = combined_2023)
summary(mode_effects_school_plus)$coefficients["as.factor(source)paper:gendermale",]

# Satisfaction with friends
mode_effects_school_plus <- lm(satis_friends ~ as.factor(source)*gender + gender*as.factor(school_id) + gender*age, data = combined_2023)
summary(mode_effects_school_plus)$coefficients["as.factor(source)paper:gendermale",]

# Satisfaction with mom
mode_effects_school_plus <- lm(satis_mom ~ as.factor(source)*gender + gender*as.factor(school_id) + gender*age, data = combined_2023)
summary(mode_effects_school_plus)$coefficients["as.factor(source)paper:gendermale",]

# Satisfaction with dad
mode_effects_school_plus <- lm(satis_dad ~ as.factor(source)*gender + gender*as.factor(school_id) + gender*age, data = combined_2023)
summary(mode_effects_school_plus)$coefficients["as.factor(source)paper:gendermale",]

# Satisfaction with leisure
mode_effects_school_plus <- lm(satis_leisure ~ as.factor(source)*gender + gender*as.factor(school_id) + gender*age, data = combined_2023)
summary(mode_effects_school_plus)$coefficients["as.factor(source)paper:gendermale",]

# Satisfaction with dwelling
mode_effects_school_plus <- lm(satis_dwell ~ as.factor(source)*gender + gender*as.factor(school_id) + gender*age, data = combined_2023)
summary(mode_effects_school_plus)$coefficients["as.factor(source)paper:gendermale",]

# Satisfaction with grades
mode_effects_school_plus <- lm(satis_grades ~ as.factor(source)*gender + gender*as.factor(school_id) + gender*age, data = combined_2023)
summary(mode_effects_school_plus)$coefficients["as.factor(source)paper:gendermale",]

```

## What happens to the explanandum if we take into account source
```{r }
# The original
simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender), data = combined, weights = wph)
summary(simple_model)

# Adding source
simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender) + as.factor(source)*as.factor(gender), data = combined, weights = wph)
summary(simple_model)

# the effect does not change sign but is reduced to non-significant .16

# Adding additional controls
simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender) + as.factor(source)*as.factor(gender) + as.factor(school_id), data = combined, weights = wph)
summary(simple_model)


```


## Just school fixed-effects for the explanandum
So, a question that arises here is: What's up with assessment mode?
Girls might score lower on satisfaction on the tablet because the school which use tablets are subtly different in a manner that explains larger gender gaps.
Or, alternatively, girls might score lower on satisfaction on tablet because of the tablet.
In the former case, assuming that this completely drives the explanandum, controlling for the school_id should make the explanandum disappear (regardless of source).

```{r }
simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender) + as.factor(school_id)*as.factor(gender), data = combined, weights = wph)
summary(simple_model)
# But this is still barely hanging in there.
# So between-school heterogeneity potentially accounts for some of the observed differences
# but it cannot fully explain away the pattern, boys are still 0.2 SD happier considering the differene in difference

```


## What happens if we take into account both migration status and source in 2023

```{r }
# Lets split things up -- do we get the same pattern if we split by language at home?

mode_effects_language <- lm(satis ~ as.factor(source)*gender*as.factor(german_at_home), data = combined_2023)
summary(mode_effects_language)

# there is a weak interaction here
pred_mode_effects_language <- predictions(mode_effects_language,
                 by = c("gender", "german_at_home", "source"))
pred_mode_effects_language

# let us do a plot for this

plot_frame <- pred_mode_effects_language %>%
  mutate(
    group = interaction(german_at_home, gender, source),
    group = factor(group, levels = c(
      "1.female.paper", "1.female.online",
      "1.male.paper", "1.male.online",
      "0.female.paper", "0.female.online",
      "0.male.paper", "0.male.online"
    ))
  )

ggplot(plot_frame, aes(x = group, y = estimate, fill = source)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  geom_errorbar(
    aes(ymin = conf.low, ymax = conf.high),
    width = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  labs(
    x = "Groups",
    y = "Estimate",
    fill = "Source"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```
```{r }
# If we put this together
# Can we get rid of the explanandum?
simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender) + as.factor(source)*as.factor(gender), data = combined[combined$german_at_home == 1,], weights = wph)
summary(simple_model) # no more difference to be explained

simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender) + as.factor(source)*as.factor(gender) + as.factor(school_id), data = combined[combined$german_at_home == 1,], weights = wph)
summary(simple_model) # no more difference to be explained

simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender) + as.factor(source)*as.factor(gender), data = combined[combined$german_at_home == 0,], weights = wph)
summary(simple_model) # large difference here remains



simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender) + as.factor(source)*as.factor(gender) + as.factor(school_id)*as.factor(gender), data = combined[combined$german_at_home == 1,], weights = wph)
summary(simple_model) # no more difference to be explained

simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender) + as.factor(source)*as.factor(gender)+ as.factor(school_id), data = combined[combined$german_at_home == 0,], weights = wph)
summary(simple_model) # large difference here remains

simple_model <- lm(scale(satis) ~ as.factor(year)*as.factor(gender) + as.factor(source)*as.factor(gender)+ as.factor(school_id), data = combined[combined$german_at_home == 0,], weights = wph)
summary(simple_model) # large difference here remains

```
