---
title: "Check our explanation for different outcomes"
author: "Julia Rohrer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("synthetic_data")
combined <- syn_dat
library(haven)
library(marginaleffects)
library(ggplot2)
library(patchwork)
combined$source <- as.factor(combined$source)
```

## Check other outcomes
Let's see whether our favored explanation also holds up for the other satisfaction variables


### Satisfaction with the amount of money with which you have to get by (financial satisfaction)
```{r satis_money}
# Fit the model
explanandum_paper_only_money <- lm(satis_money ~ as.factor(year) + gender + as.factor(year):gender +
                                 as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                                 as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                 as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang), 
                  data = combined[combined$source == "paper",])

# Demographics like in 2023
temp1 <- combined[combined$year == 2023,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023,]
temp3$year <- 2023
all2023 <- rbind(temp1, temp2, temp3)

# Predictions
pred <- predictions(explanandum_paper_only_money,
                 by = c("gender", "year"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)

# Gender gap by year
comps <- avg_comparisons(explanandum_paper_only_money, variables = "gender", by = "year",
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps)

# Test biggest contrast: 2010 versus 2023
max_comp <- avg_comparisons(explanandum_paper_only_money, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2023,
                vcov = ~ unique_classroom)
print(max_comp)

# Gender gap by year and migration background
comps_by_migback <- avg_comparisons(explanandum_paper_only_money, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps_by_migback)

# Widening of the gap for no migration background
# 2023
print(avg_comparisons(explanandum_paper_only_money, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))

# Widening of the gap for migration background
# 2023
print(avg_comparisons(explanandum_paper_only_money, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))

pred <- avg_predictions(explanandum_paper_only_money,
                        newdata = all2023,
                        by = c("gender", "year", "mig_lang"),
                        vcov = ~ unique_classroom)

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean financial satisfaction") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean financial satisfaction") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
#ggsave("Plots/paper_satis_money.png", width = 6, height = 3)

```



### Satisfaction with your relationships to friends

```{r satis_friends}
# Fit the model
explanandum_paper_only_friends <- lm(satis_friends ~ as.factor(year) + gender + as.factor(year):gender +
                                 as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                                 as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                 as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang), 
                  data = combined[combined$source == "paper",])

# Demographics like in 2023
temp1 <- combined[combined$year == 2023,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023,]
temp3$year <- 2023
all2023 <- rbind(temp1, temp2, temp3)

# Predictions
pred <- predictions(explanandum_paper_only_friends,
                 by = c("gender", "year"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)

# Gender gap by year
comps <- avg_comparisons(explanandum_paper_only_friends, variables = "gender", by = "year",
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps)

# Test biggest contrast: 2010 versus 2023
max_comp <- avg_comparisons(explanandum_paper_only_friends, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2023,
                vcov = ~ unique_classroom)
print(max_comp)

# Gender gap by year and migration background
comps_by_migback <- avg_comparisons(explanandum_paper_only_friends, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps_by_migback)

# Widening of the gap for no migration background
# 2023
print(avg_comparisons(explanandum_paper_only_friends, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))

# Widening of the gap for migration background
# 2023
print(avg_comparisons(explanandum_paper_only_friends, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))

pred <- avg_predictions(explanandum_paper_only_friends,
                        newdata = all2023,
                        by = c("gender", "year", "mig_lang"),
                        vcov = ~ unique_classroom)

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean friendship satisfaction") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean friendship satisfaction") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
#ggsave("Plots/paper_satis_friends.png", width = 6, height = 3)

```


### Satisfaction with your relationship to your mother

```{r mom}
# Fit the model
explanandum_paper_only_mom <- lm(satis_mom ~ as.factor(year) + gender + as.factor(year):gender +
                                 as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                                 as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                 as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang), 
                  data = combined[combined$source == "paper",])

# Demographics like in 2023
temp1 <- combined[combined$year == 2023,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023,]
temp3$year <- 2023
all2023 <- rbind(temp1, temp2, temp3)

# Predictions
pred <- predictions(explanandum_paper_only_mom,
                 by = c("gender", "year"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)

# Gender gap by year
comps <- avg_comparisons(explanandum_paper_only_mom, variables = "gender", by = "year",
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps)

# Test biggest contrast: 2010 versus 2023
max_comp <- avg_comparisons(explanandum_paper_only_mom, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2023,
                vcov = ~ unique_classroom)
print(max_comp)

# Gender gap by year and migration background
comps_by_migback <- avg_comparisons(explanandum_paper_only_mom, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps_by_migback)

# Widening of the gap for no migration background
# 2023
print(avg_comparisons(explanandum_paper_only_mom, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))

# Widening of the gap for migration background
# 2023
print(avg_comparisons(explanandum_paper_only_mom, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))

pred <- avg_predictions(explanandum_paper_only_mom,
                        newdata = all2023,
                        by = c("gender", "year", "mig_lang"),
                        vcov = ~ unique_classroom)

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean satisfaction w mom") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean satisfaction w mom") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
#ggsave("Plots/paper_satis_mother.png", width = 6, height = 3)

```

### Satisfaction with your relationship to your father

```{r dad}
# Fit the model
explanandum_paper_only_dad <- lm(satis_dad ~ as.factor(year) + gender + as.factor(year):gender +
                                 as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                                 as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                 as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang), 
                  data = combined[combined$source == "paper",])

# Demographics like in 2023
temp1 <- combined[combined$year == 2023,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023,]
temp3$year <- 2023
all2023 <- rbind(temp1, temp2, temp3)

# Predictions
pred <- predictions(explanandum_paper_only_dad,
                 by = c("gender", "year"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)

# Gender gap by year
comps <- avg_comparisons(explanandum_paper_only_dad, variables = "gender", by = "year",
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps)

# Test biggest contrast: 2010 versus 2023
max_comp <- avg_comparisons(explanandum_paper_only_dad, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2023,
                vcov = ~ unique_classroom)
print(max_comp)

# Gender gap by year and migration background
comps_by_migback <- avg_comparisons(explanandum_paper_only_dad, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps_by_migback)

# Widening of the gap for no migration background
# 2023
print(avg_comparisons(explanandum_paper_only_dad, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))

# Widening of the gap for migration background
# 2023
print(avg_comparisons(explanandum_paper_only_dad, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))

pred <- avg_predictions(explanandum_paper_only_dad,
                        newdata = all2023,
                        by = c("gender", "year", "mig_lang"),
                        vcov = ~ unique_classroom)

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean satisfaction w father") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean satisfaction w father") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
#ggsave("Plots/paper_satis_dad.png", width = 6, height = 3)
```

### Satisfaction with your leisure time activities

```{r leisure}
# Fit the model
explanandum_paper_only_leisure <- lm(satis_leisure ~ as.factor(year) + gender + as.factor(year):gender +
                                 as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                                 as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                 as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang), 
                  data = combined[combined$source == "paper",])

# Demographics like in 2023
temp1 <- combined[combined$year == 2023,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023,]
temp3$year <- 2023
all2023 <- rbind(temp1, temp2, temp3)

# Predictions
pred <- predictions(explanandum_paper_only_leisure,
                 by = c("gender", "year"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)

# Gender gap by year
comps <- avg_comparisons(explanandum_paper_only_leisure, variables = "gender", by = "year",
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps)

# Test biggest contrast: 2010 versus 2023
max_comp <- avg_comparisons(explanandum_paper_only_leisure, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2023,
                vcov = ~ unique_classroom)
print(max_comp)

# Gender gap by year and migration background
comps_by_migback <- avg_comparisons(explanandum_paper_only_leisure, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps_by_migback)

# Widening of the gap for no migration background
# 2023
print(avg_comparisons(explanandum_paper_only_leisure, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))

# Widening of the gap for migration background
# 2023
print(avg_comparisons(explanandum_paper_only_leisure, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))

pred <- avg_predictions(explanandum_paper_only_leisure,
                        newdata = all2023,
                        by = c("gender", "year", "mig_lang"),
                        vcov = ~ unique_classroom)

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean leisure satisfaction") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean leisure satisfaction") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
#ggsave("Plots/paper_satis_leisure.png", width = 6, height = 3)
```


### Satisfaction with your school grades

```{r grades}
# Fit the model
explanandum_paper_only_grades <- lm(satis_grades ~ as.factor(year) + gender + as.factor(year):gender +
                                 as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                                 as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                 as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang), 
                  data = combined[combined$source == "paper",])

# Demographics like in 2023
temp1 <- combined[combined$year == 2023,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023,]
temp3$year <- 2023
all2023 <- rbind(temp1, temp2, temp3)

# Predictions
pred <- predictions(explanandum_paper_only_grades,
                 by = c("gender", "year"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)

# Gender gap by year
comps <- avg_comparisons(explanandum_paper_only_grades, variables = "gender", by = "year",
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps)

# Test biggest contrast: 2010 versus 2023
max_comp <- avg_comparisons(explanandum_paper_only_grades, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2023,
                vcov = ~ unique_classroom)
print(max_comp)

# Gender gap by year and migration background
comps_by_migback <- avg_comparisons(explanandum_paper_only_grades, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps_by_migback)

# Widening of the gap for no migration background
# 2023
print(avg_comparisons(explanandum_paper_only_grades, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))

# Widening of the gap for migration background
# 2023
print(avg_comparisons(explanandum_paper_only_grades, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))

pred <- avg_predictions(explanandum_paper_only_grades,
                        newdata = all2023,
                        by = c("gender", "year", "mig_lang"),
                        vcov = ~ unique_classroom)

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean school grades satisfaction") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean school grades satisfaction") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
#ggsave("Plots/paper_satis_grades.png", width = 6, height = 3)
```

### Satisfaction with your housing situation

```{r dwell}
# Fit the model
explanandum_paper_only_dwell <- lm(satis_dwell ~ as.factor(year) + gender + as.factor(year):gender +
                                 as.factor(age) + as.factor(year):as.factor(age) + gender:as.factor(age) + as.factor(year):gender:as.factor(age) +
                                 as.factor(schooltype) + as.factor(year):as.factor(schooltype) + gender:as.factor(schooltype) + as.factor(year):gender:as.factor(schooltype) +
                                 as.factor(mig_lang) + as.factor(year):as.factor(mig_lang) + gender:as.factor(mig_lang) + as.factor(year):gender:as.factor(mig_lang), 
                  data = combined[combined$source == "paper",])

# Demographics like in 2023
temp1 <- combined[combined$year == 2023,]
temp1$year <- 2010
temp2 <- combined[combined$year == 2023,]
temp2$year <- 2015
temp3 <- combined[combined$year == 2023,]
temp3$year <- 2023
all2023 <- rbind(temp1, temp2, temp3)

# Predictions
pred <- predictions(explanandum_paper_only_dwell,
                 by = c("gender", "year"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)

# Gender gap by year
comps <- avg_comparisons(explanandum_paper_only_dwell, variables = "gender", by = "year",
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps)

# Test biggest contrast: 2010 versus 2023
max_comp <- avg_comparisons(explanandum_paper_only_dwell, variables = "gender", by = "year",
                hypothesis = "b1 = b3",
                 newdata = all2023,
                vcov = ~ unique_classroom)
print(max_comp)

# Gender gap by year and migration background
comps_by_migback <- avg_comparisons(explanandum_paper_only_dwell, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom)
print(comps_by_migback)

# Widening of the gap for no migration background
# 2023
print(avg_comparisons(explanandum_paper_only_dwell, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
      hypothesis = "b1 = b5"))

# Widening of the gap for migration background
# 2023
print(avg_comparisons(explanandum_paper_only_dwell, variables = "gender", by = c("year", "mig_lang"),
                 newdata = all2023,
                 vcov = ~ unique_classroom,
                 hypothesis = "b2 = b6"))

pred <- avg_predictions(explanandum_paper_only_dwell,
                        newdata = all2023,
                        by = c("gender", "year", "mig_lang"),
                        vcov = ~ unique_classroom)

# Plot the results
# No migration background
cat_mig_lang_0 <- ggplot(pred[pred$mig_lang == 0,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("No migration background") +
  ylab("Mean housing satisfaction") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  xlab("Year") +
  labs(color = "Gender")

# With migration background
cat_mig_lang_1 <- ggplot(pred[pred$mig_lang == 1,], aes(x = year, group = gender, 
                                 color = gender,
                                 y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_line(position = position_dodge(width = .5)) +
  geom_errorbar(width = .5, position = position_dodge(width = .5)) +
  coord_cartesian(ylim = c(mean(combined$satis) - sd(combined$satis), mean(combined$satis) + sd(combined$satis))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Migration background") +
  scale_x_continuous(breaks = c(2010, 2015, 2023)) +
  ylab("Mean housing satisfaction") +
  xlab("Year") +
  labs(color = "Gender")

# Combine into single plot
combined_cat_mig_lang <- (cat_mig_lang_0|cat_mig_lang_1) + plot_layout(guides = "collect")
combined_cat_mig_lang
#ggsave("Plots/paper_satis_dwell.png", width = 6, height = 3)
```